---
title: "Okinawa Bulk Coral Physiology 2025"
author: "Maya Powell"
date-created: "November 11th 2025"
date-modified: today
format:
  html:
    toc: true
#    theme: darkly
editor: visual
warning: false
---

## Setup

Load libraries

```{r}
library(tidyverse)
library(here)
library(dplyr)
library(car)
library(PNWColors)
library(performance)
library(ggpubr)
```

Assign species colors for all plots

```{r}
sp_cols <- c(
  "Acropora hyacinthus" = '#d8aedd',
  "Echinopora lamellosa" = '#ba7999',
  "Favites complanata"   = '#dd4124',
  "Montipora aequituberculata" = '#ed8b00',
  "Montipora vietnamensis" = '#efbc82',
  "Pachyseris rugosa" = '#edd746',
  "Pocillopora eydouxi" = '#d0e2af',
  "Porites cylindrica" = '#45681e',
  "Porites rus" = '#7bbcd5',
  "Turbinaria frondens" = '#00496f'
)
```

Function to calculate SE throughout

```{r}
se_fun <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) return(NA_real_)
  sd(x, na.rm = TRUE) / sqrt(n)
}
```

## Dry weight

Load data

```{r}
avg_DW <- read.csv(here("Data", "Physiology", "Average_Dry_Weight.csv"))
avg_DW <- avg_DW %>% rename(full_species = species_long)
```

```{r}
#code to calculate average dry weight per fragment
#needed initially but not after writing average dry weight datasheet
# #read in metadata
# phys_meta <- read.csv(here("Data", "Physiology", "Physio_meta.csv"))
# 
# DW <- read.csv(here("Data", "Physiology", "Dry_Weight_Oki2025.csv"))
# 
# DW <- DW %>% left_join(phys_meta, by = "frag_ID")
# 
# #calculate dry weight
# DW <- DW %>%
#   mutate(dw_g_mL = dry_pan - pan_weight,
#          dw_total_g = dw_g_mL*blastate_vol_mL,
#          dw_g_cm2 = dw_total_g/SA_cm2,
#          dw_mg_cm2 = dw_g_cm2*1000)
# 
# #average across the reps
# avg_DW <- DW %>%
#   group_by(frag_ID) %>%
#   summarise(dw_g_mL = mean(dw_g_mL),
#             dw_total_g = mean(dw_total_g),
#             dw_g_cm2 = mean(dw_g_cm2),
#             dw_mg_cm2 = mean(dw_mg_cm2)) %>%
#   left_join(phys_meta, by = "frag_ID")
# 
# write.csv(avg_DW, here("Data", "Physiology", "Average_Dry_Weight.csv"), row.names = FALSE)
```

Generate dry weight summary data

```{r}
dw_summary <- avg_DW %>%
  group_by(full_species) %>%
  summarise(
    n    = sum(!is.na(dw_mg_cm2)),
    mean = mean(dw_mg_cm2, na.rm = TRUE),
    se   = se_fun(dw_mg_cm2),
    .groups = "drop")
```

See if there is a difference between species

```{r}
DW.mod.spp <- lm(dw_mg_cm2~full_species, data = avg_DW)
Anova(DW.mod.spp)
summary(DW.mod.spp)
check_model(DW.mod.spp)
```

Emmeans comparisons

```{r}
emm_obj <- emmeans::emmeans(DW.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#a, a, a, a, ab, ab, b, b
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols, ) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","dw_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder based on mean values and then plot dry weight

```{r}
means <- avg_DW %>% group_by(species) %>% summarize(m = mean(dw_mg_cm2, na.rm = TRUE), .groups = "drop")

avg_DW_ordered <- avg_DW %>% left_join(dw_summary, by = "full_species") %>% mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

dw_plot <- ggplot() +
  geom_jitter(data = avg_DW_ordered, aes(x = full_species, y = dw_mg_cm2, color = full_species), width = 0.15, alpha = 0.8) +
  stat_summary(data = avg_DW_ordered, aes(x = full_species, y = dw_mg_cm2), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "a", "a", "a", "ab", "ab", "b"))+
  geom_errorbar(data = dw_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = dw_summary, aes(x = full_species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  ylim(25,325)+
  labs(x = "Species", color = "Species",
       y = expression("Tissue biomass" ~ (mg ~ cm^{-2})))
dw_plot

#ggsave(here("Output", "Physiology", "dryweight_species_jitter.pdf"), dw_plot, h = 8, w = 10)
```

## Chlorophyll a

Initial processing - only needed once

```{r}
#read in data
#data is corrected for blank rates and surface area and blanks are removed
#uncomment this section if you need to redo prep for chla dataframe
# chla_plate1 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate1_summary.csv"))
# chla_plate2 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate2_summary.csv"))
# chla_plate3 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate3_summary.csv"))
# chla_plate4 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate4_summary.csv"))
# 
# # concatenate all plates
# chla_all_corr <- bind_rows(chla_plate1,chla_plate2,chla_plate3,chla_plate4)
# chla_all_corr <- chla_all_corr %>% rename(frag_ID = Sample.Name)
# 
# #add metadata, SA, and symbiont density to calc chla/sym
# SA_frag_final <- read.csv(here("Data","SurfaceArea","SA_fragments_final.csv"))
# chla_meta <- SA_frag_final %>% filter(Frag_type == "physio") %>% 
#   select(frag_ID, full_species, species, calc_SA)
# avg_sym <- read.csv(here("Data", "Physiology", "Average_Sym_Density.csv"))
# sym_select <- avg_sym %>% 
#   select(frag_ID, sym_total)
# 
# chla_data <- left_join(chla_all_corr, chla_meta, by = "frag_ID") %>%
#   rename(SA_cm2 = calc_SA)
# chla_data <- left_join(chla_data, sym_select, by = "frag_ID") #only have 7 species here for sym so will have to subset
# 
# #save data
# write_csv(chla_data, here("Data", "Physiology", "Chla_avg.csv"))
```

Read in data

```{r}
chla_avg <- read.csv(here("Data", "Physiology", "Chla_avg.csv"))

chla_avg <- chla_avg %>% 
  #filter(full_species %in% sp_keep) %>% 
  filter(frag_ID != "C07") %>% #remove outlier frag
  mutate(chla_ug_total = (chla_mean*50)) %>% #multiply by slurry volume to get total chla
  mutate(chla_pg_sym = (chla_ug_total/sym_total)*1000000) #divide by total symbionts and multiply by 1000000 to get pg

```

#### Chla/cm2

Make chla summary data

```{r}
#summarize chla/cm2 by species
chla_summary <- chla_avg %>%
  group_by(full_species) %>%
  summarise(n = sum(!is.na(chla_ug_cm2_mean)),
            mean = mean(chla_ug_cm2_mean, na.rm = TRUE),
            se = se_fun(chla_ug_cm2_mean),
            .groups = "drop")
```

See if there's a difference between species for chla

```{r}
chla.mod.spp <- lm(chla_ug_cm2_mean~full_species, data = chla_avg)
Anova(chla.mod.spp)
summary(chla.mod.spp)
check_model(chla.mod.spp)
```

Generate emmeans comparisons and plot

```{r}
emm_obj <- emmeans::emmeans(chla.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#a, a, a, a, a, a, ab, ab, b, b
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols, ) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","chla_cm2_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Order and plot data by species

```{r}
#reorder based on mean values
means <- chla_avg %>% group_by(species) %>% summarize(m = mean(chla_ug_cm2_mean, na.rm = TRUE), .groups = "drop")

chla_avg_ordered <- chla_avg %>% 
  left_join(chla_summary, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

chla_plot <- ggplot() +
  geom_jitter(data = chla_avg_ordered, aes(x = full_species, y = chla_ug_cm2_mean, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = chla_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6)+
  geom_point(data = chla_summary, aes(x = full_species, y = mean), size = 2) +
  stat_summary(data = chla_avg_ordered, aes(x = full_species, y = chla_ug_cm2_mean), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "a", "a", "a", "a", "a", "ab", "ab", "b", "b"))+
  theme_bw(base_size = 22)+
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  labs(x = "Species", color = "Species", y = expression("Chlorophyll a content" ~ (mu*g ~ cm^{-2}))) +
  ylim(0,22)
chla_plot

#ggsave(here("Output", "Physiology", "chla_cm2_species_jitter.pdf"), chla_plot, h = 8, w = 10)
```

#### Chla/symbiont

Only seven species for sym data for now so edit dataframe accordingly

```{r}
sp_keep <- c("Acropora hyacinthus","Echinopora lamellosa", "Favites complanata","Montipora aequituberculata","Pocillopora eydouxi","Porites rus","Turbinaria frondens")
chla_7sp <- chla_avg %>% filter(full_species %in% sp_keep)

```

Summarize

```{r}
chla_sym_summary <- chla_7sp %>%
  group_by(full_species) %>%
  summarise(n = sum(!is.na(chla_pg_sym)),
            mean = mean(chla_pg_sym, na.rm = TRUE),
            se = se_fun(chla_pg_sym),
            .groups = "drop")
```

See if there's a difference between species

```{r}
chla.sym.spp <- lm(chla_pg_sym~full_species, data = chla_7sp)
Anova(chla.sym.spp)
summary(chla.sym.spp)
check_model(chla.sym.spp)
```

Emmeans comparison and plot

```{r}
emm_obj <- emmeans::emmeans(chla.sym.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#a, ab, ab, abc, abc, bc, c
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","chla_sym_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder and plot by species

```{r}
#reorder based on mean values
means <- chla_7sp %>% group_by(species) %>% summarize(m = mean(chla_pg_sym, na.rm = TRUE), .groups = "drop")

chla_7sp_ordered <- chla_7sp %>% 
  left_join(chla_sym_summary, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

chla_sym_plot <- ggplot() +
  geom_jitter(data = chla_7sp_ordered, aes(x = full_species, y = chla_pg_sym, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = chla_sym_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6)+
  geom_point(data = chla_sym_summary, aes(x = full_species, y = mean), size = 2) +
  stat_summary(data = chla_7sp_ordered, aes(x = full_species, y = chla_pg_sym), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "ab", "ab", "abc", "abc", "bc", "c"))+
  theme_bw(base_size = 22)+
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  labs(x = "Species", color = "Species", y = expression("Chlorophyll a content" ~ (pg ~ symbiont^{-2}))) +
  ylim(0,11)
chla_sym_plot

ggsave(here("Output", "Physiology", "chla_sym_species_jitter.pdf"), chla_sym_plot, h = 8, w = 10)
```

## Symbiont density

Load data

```{r}
avg_sym <- read.csv(here("Data", "Physiology", "Average_Sym_Density.csv"))
avg_sym <- avg_sym %>% 
  filter(frag_ID != "C07") %>% 
  filter(frag_ID != "D10") #remove crazy outliers!
```

Needed initially to calculate avg per frag sym density

```{r}
#code to calculate average Symbiont Density per fragment
#needed initially but not after writing average Symbiont Density datasheet
# phys_meta <- read.csv(here("Data", "Physiology", "Physio_meta.csv"))
# 
# #read in sym photo id sheet and sym counts
# counts <- read.csv(here("Data", "Physiology", "sym_counts_all.csv"))
# photos <- read.csv(here("Data", "Physiology", "sym_photos_all.csv"))
# 
# #combine
# sym_raw <- photos %>% left_join(counts, by = "image_id")
# 
# #add together counts from photos from 2 focus settings for photos
# img_rm <- c("P8184974","P8184975","P8184976","P8174396","P8174397","P8174662","P8174663","P8174668","P8174669")
# sym_add <- sym_raw %>%
#   group_by(frag_ID,rep) %>% mutate(count_2 = sum(count)) %>%
#   filter(image_no == 1) %>% #keep only 1 row of data for each
#   filter(!image_id %in% img_rm) #filter out images to remove based on SE >15% of mean (as calculated below)
# 
# #average sym counts across 6 reps for each id
# #and calculate SE
# sym_avg <- sym_add %>%
#   group_by(frag_ID) %>% mutate(count_avg = mean(count_2)) %>%
#   mutate(count_se = sd(count_2)/sqrt(length(count_2))) %>%
#   #check to see which have highest error and look at add data for those >15%
#   mutate(count_var = count_se/count_avg) %>%
#   filter(rep == 6) #keep only one row of data for each
# 
# #error >0.15 = C02, E04, H08
# #C02 - one data point is 6 while the others are >50, likely remove - check photos
# #E04 - two counts of 9 and 6, all others >30, likely remove - check photos
# #H08 - two counts in the 300s while all others in the 100s, likely remove - check photos
# 
# #now calculate actual symbiont density
# #the 1x1mm grid should be 0.1uL
# #we want cells/mL so 10,000 x our counts (1000uL in a mL/0.1uL = x 10,000)
# #homogenate volume for all is 50mL
# #SA is calculated as well here
# #combine with metadata for calculations
# 
# sym <- sym_avg %>%
#   select(frag_ID, count_avg) %>% #keep only relevant columns
#   left_join(phys_meta, by = "frag_ID") %>% #add metadata
#   mutate(sym_mL = count_avg*10000) %>%
#   mutate(sym_total = sym_mL*blastate_vol_mL) %>%
#   mutate(sym_cm2 = sym_total/SA_cm2)
# 
# write.csv(sym, here("Data", "Physiology", "Average_Sym_Density.csv"), row.names = FALSE)
```

Summary data

```{r}
sym_summary <- avg_sym %>%
  group_by(full_species) %>%
  summarise(
    n    = sum(!is.na(sym_cm2)),
    mean = mean(sym_cm2, na.rm = TRUE),
    se   = se_fun(sym_cm2),
    .groups = "drop")

```

See if there's a difference between species

```{r}
sym.mod.spp <- lm(sym_cm2~full_species, data = avg_sym)
Anova(sym.mod.spp)
summary(sym.mod.spp)
check_model(sym.mod.spp)
```

Emmeans comparison and plot

```{r}
emm_obj <- emmeans::emmeans(sym.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#a, ab, ab, b, b, bc, c
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","sym_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder and plot by species

```{r}
#reorder based on mean values
means <- avg_sym %>% group_by(species) %>% summarize(m = mean(sym_cm2, na.rm = TRUE), .groups = "drop")

avg_sym_ordered <- avg_sym %>% left_join(sym_summary, by = "full_species") %>% mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

sym_plot <- ggplot() +
  geom_jitter(data = avg_sym_ordered, aes(x = full_species, y = sym_cm2, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = sym_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = sym_summary, aes(x = full_species, y = mean), size = 2) +
  stat_summary(data = avg_sym_ordered, aes(x = full_species, y = sym_cm2), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "ab", "ab", "b", "b", "bc", "c"))+
  theme_bw(base_size = 22) +
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  #ylim(25,325)+
  labs(x = "Species", color = "Species",
       y = expression("Symbiont Density" ~ (cells ~ cm^{-2})))
sym_plot

ggsave(here("Output", "Physiology", "sym_species_jitter.pdf"), sym_plot, h = 8, w = 10)
```

## Protein

Load data

```{r}
prot <- read.csv(here("Data", "Physiology", "protein_all_summary.csv"))
```

Protein summary data

```{r}
prot_summary <- prot %>%
  group_by(full_species) %>%
  summarise(
    n    = sum(!is.na(prot_ug_cm2)),
    mean = mean(prot_ug_cm2, na.rm = TRUE),
    se   = se_fun(prot_ug_cm2),
    .groups = "drop")
```

See if there's a difference between species

```{r}
prot.mod.spp <- lm(prot_ug_cm2~full_species, data = prot)
Anova(prot.mod.spp)
summary(prot.mod.spp)
check_model(prot.mod.spp)
```

Emmeans comparisons and plot

```{r}
emm_obj <- emmeans::emmeans(prot.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols, ) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output","Physiology","prot_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder and plot species comparisons

```{r}
#reorder based on mean values
means <- prot %>% group_by(species) %>% summarize(m = mean(prot_ug_cm2, na.rm = TRUE), .groups = "drop")

prot_ordered <- prot %>% left_join(prot_summary, by = "full_species") %>% mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

prot_plot <- ggplot() +
  geom_jitter(data = prot_ordered, aes(x = full_species, y = prot_ug_cm2, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = prot_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = prot_summary, aes(x = full_species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  stat_summary(data = prot_ordered, aes(x = full_species, y = prot_ug_cm2), geom = "text", fun = max, vjust = -0.5, size = 8,
    label = c("a", "a", "ab", "ab", "a", "a", "b", "ab", "ab", "ab"))+
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  ylim(0,4100)+
  labs(x = "Species", color = "Species",
       y = expression("Protein Content" ~ (mu*g ~ cm^{-2})))
prot_plot

#ggsave(here("Output", "Physiology", "prot_species_jitter.pdf"), prot_plot, h = 8, w = 10)

```

## All plots together

Put all physio plots together

```{r}
physio_plots <- ggarrange(chla_plot, sym_plot, dw_plot, prot_plot, nrow = 2, ncol = 2, legend = "right", common.legend = TRUE)
physio_plots

#ggsave(here("Output","Physiology","physio_plots.pdf"), physio_plots, h = 10, w = 15)
```
