---
title: "Okinawa Bulk Coral Physiology 2025"
author: "Maya Powell"
date-created: "November 11th 2025"
date-modified: today
format:
  html:
    toc: true
#    theme: darkly
editor: visual
warning: false
---

## Setup

Load libraries

```{r}
#install packages as needed
library(tidyverse)
library(here)
library(dplyr)
library(car)
library(PNWColors)
library(see)
library(performance)
library(emmeans)
library(ggplot2)
library(ggpubr)
```

Assign species colors for all plots

```{r}
sp_cols <- c(
  "Acropora hyacinthus" = '#d8aedd',
  "Echinopora lamellosa" = '#ba7999',
  "Favites complanata"   = '#dd4124',
  "Montipora aequituberculata" = '#ed8b00',
  "Montipora vietnamensis" = '#efbc82',
  "Pachyseris rugosa" = '#edd746',
  "Pocillopora eydouxi" = '#d0e2af',
  "Porites cylindrica" = '#45681e',
  "Porites rus" = '#7bbcd5',
  "Turbinaria frondens" = '#00496f'
)
```

Function to calculate SE throughout

```{r}
se_fun <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) return(NA_real_)
  sd(x, na.rm = TRUE) / sqrt(n)
}
```

## Dry weight

Load data

```{r}
avg_DW <- read.csv(here("Data", "Physiology", "Average_Dry_Weight.csv"))
avg_DW <- avg_DW %>% rename(full_species = species_long)
```

```{r}
#code to calculate average dry weight per fragment
#needed initially but not after writing average dry weight datasheet
#read in metadata
phys_meta <- read.csv(here("Data", "Physiology", "Physio_meta_all.csv"))

DW <- read.csv(here("Data", "Physiology", "afdw_sesoko.csv"))

DW <- DW %>% left_join(phys_meta, by = "frag_ID")

#calculate dry weight & ash free dry weight
DW <- DW %>%
  mutate(dw_g_mL = dry_pan - pan_weight,
         dw_total_g = dw_g_mL*blastate_vol_mL,
         dw_g_cm2 = dw_total_g/SA_cm2,
         dw_mg_cm2 = dw_g_cm2*1000,
         afdw_g_mL = dry_pan - ash_pan,
         afdw_total_g = afdw_g_mL*blastate_vol_mL,
         afdw_g_cm2 = afdw_total_g/SA_cm2,
         afdw_mg_cm2 = afdw_g_cm2*1000)

#average across the reps
#use all 6 reps across two years for dry weight
avg_DW <- DW %>%
  group_by(frag_ID) %>%
  dplyr::select(-ash_pan,-AFDW,-afdw_g_mL,-afdw_total_g,-afdw_g_cm2,-afdw_mg_cm2) %>%
  summarise(dw_g_mL = mean(dw_g_mL),
            dw_total_g = mean(dw_total_g),
            dw_g_cm2 = mean(dw_g_cm2),
            dw_mg_cm2 = mean(dw_mg_cm2))
  
#use only the 3 reps over one year for AFDW
avg_AFDW <- DW %>%
  drop_na(afdw_g_cm2) %>%
  group_by(frag_ID) %>%
  summarise(afdw_g_mL = mean(afdw_g_mL),
            afdw_total_g = mean(afdw_total_g),
            afdw_g_cm2 = mean(afdw_g_cm2),
            afdw_mg_cm2 = mean(afdw_mg_cm2)) %>%
  left_join(avg_DW, by = "frag_ID") %>%
  left_join(phys_meta, by = "frag_ID")

write.csv(avg_AFDW, here("Data", "Physiology", "Average_Ash_Free)Dry_Weight.csv"), row.names = FALSE)
```

Generate dry weight summary data

```{r}
afdw_summary <- avg_AFDW %>%
  group_by(full_species) %>%
  summarise(
    n_dw = sum(!is.na(dw_mg_cm2)),
    mean_dw = mean(dw_mg_cm2, na.rm = TRUE),
    se_dw = se_fun(dw_mg_cm2),
    n_afdw = sum(!is.na(afdw_mg_cm2)),
    mean_afdw = mean(afdw_mg_cm2, na.rm = TRUE),
    se_afdw = se_fun(afdw_mg_cm2))

```

See if there is a difference between species

```{r}
#dry weight model
DW.mod.spp <- lm(dw_mg_cm2~full_species, data = avg_AFDW)
Anova(DW.mod.spp)
summary(DW.mod.spp)
check_model(DW.mod.spp)

#ash free dry weight model
AFDW.mod.spp <- lm(afdw_mg_cm2~full_species, data = avg_AFDW)
Anova(AFDW.mod.spp)
summary(AFDW.mod.spp)
check_model(AFDW.mod.spp)
```

Emmeans comparisons - ash free dry weight

```{r}
emm_obj <- emmeans::emmeans(AFDW.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#a, a, a, a, ab, ab, b, b
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols, ) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","afdw_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder based on mean values and then plot dry weight

```{r}
#| fig.width: 10
#| fig.height: 8
means <- avg_AFDW %>% group_by(full_species) %>% summarize(m = mean(afdw_mg_cm2, na.rm = TRUE), .groups = "drop")

avg_AFDW_ordered <- avg_AFDW %>% left_join(afdw_summary, by = "full_species") %>% mutate(full_species = fct_reorder(full_species, mean_afdw))  # ascending by mean

afdw_plot <- ggplot() +
  geom_jitter(data = avg_AFDW_ordered, aes(x = full_species, y = afdw_mg_cm2, color = full_species), width = 0.15, alpha = 0.8) +
  stat_summary(data = avg_AFDW_ordered, aes(x = full_species, y = afdw_mg_cm2), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "a", "a", "a", "ab", "ab", "abc", "abc", "bc", "c"))+
  geom_errorbar(data = afdw_summary, aes(x = full_species, ymin = mean_afdw - se_afdw, ymax = mean_afdw + se_afdw), width = 0.2, linewidth = 0.6) +
  geom_point(data = afdw_summary, aes(x = full_species, y = mean_afdw), size = 2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  ylim(10,75)+
  labs(x = "Species", color = "Species",
       y = expression("Tissue biomass" ~ (mg ~ cm^{-2})))
afdw_plot

ggsave(here("Output", "Physiology", "ashfreedryweight_species_jitter.pdf"), afdw_plot, h = 5, w = 10)
```

## Chlorophyll a

Initial processing - only needed once

```{r}
#read in data
#data is corrected for blank rates and surface area and blanks are removed
#uncomment this section if you need to redo prep for chla dataframe
# chla_plate1 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate1_summary.csv"))
# chla_plate2 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate2_summary.csv"))
# chla_plate3 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate3_summary.csv"))
# chla_plate4 <- read.csv(here("Data", "ChlaSpec", "oki_chla_plate4_summary.csv"))
# 
# # concatenate all plates
# chla_all_corr <- bind_rows(chla_plate1,chla_plate2,chla_plate3,chla_plate4)
# chla_all_corr <- chla_all_corr %>% rename(frag_ID = Sample.Name)
# 
# #add metadata, SA, and symbiont density to calc chla/sym
# SA_frag_final <- read.csv(here("Data","SurfaceArea","SA_fragments_final.csv"))
# chla_meta <- SA_frag_final %>% filter(Frag_type == "physio") %>%
#   select(frag_ID, full_species, species, calc_SA)
# avg_sym <- read.csv(here("Data", "Physiology", "Average_Sym_Density.csv"))
# sym_select <- avg_sym %>%
#   select(frag_ID, sym_total)
# 
# chla_data <- left_join(chla_all_corr, chla_meta, by = "frag_ID") %>%
#   rename(SA_cm2 = calc_SA)
# chla_data <- left_join(chla_data, sym_select, by = "frag_ID") #only have 7 species here for sym so will have to subset
# 
# #save data
# write_csv(chla_data, here("Data", "Physiology", "Chla_avg.csv"))
```

Read in data

```{r}
chla_avg <- read.csv(here("Data", "Physiology", "Chla_avg.csv"))

chla_avg <- chla_avg %>% 
  #filter(full_species %in% sp_keep) %>% 
  filter(frag_ID != "E04") %>% #remove outlier frag
  mutate(chla_ug_total = (chla_mean*50)) %>% #multiply by slurry volume to get total chla
  mutate(chla_pg_sym = (chla_ug_total/sym_total)*1000000) #divide by total symbionts and multiply by 1000000 to get pg

```

#### Chla/cm2

Make chla summary data

```{r}
#summarize chla/cm2 by species
chla_summary <- chla_avg %>%
  group_by(full_species) %>%
  summarise(n = sum(!is.na(chla_ug_cm2_mean)),
            mean = mean(chla_ug_cm2_mean, na.rm = TRUE),
            se = se_fun(chla_ug_cm2_mean),
            .groups = "drop")
```

See if there's a difference between species for chla

```{r}
chla.mod.spp <- lm(chla_ug_cm2_mean~full_species, data = chla_avg)
Anova(chla.mod.spp)
summary(chla.mod.spp)
check_model(chla.mod.spp)
```

Generate emmeans comparisons and plot

```{r}
emm_obj <- emmeans::emmeans(chla.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#"a", "a", "a", "a", "a", "ab", "ab", "ab", "b", "b"
#Acropora hyacinthus - Porites cylindrica             -6.8720 1.84 90  -3.726  0.0119
#Acropora hyacinthus - Porites rus                    -7.4207 1.84 90  -4.024  0.0045
#Echinopora lamellosa - Porites cylindrica            -6.4047 1.84 90  -3.473  0.0260
#Echinopora lamellosa - Porites rus                   -6.9533 1.84 90  -3.770  0.0104
#Montipora aequituberculata - Porites cylindrica      -6.9516 1.84 90  -3.769  0.0104
#Montipora aequituberculata - Porites rus             -7.5002 1.84 90  -4.067  0.0039
#Pachyseris rugosa - Porites cylindrica               -5.7105 1.84 90  -3.096  0.0742
#Pachyseris rugosa - Porites rus                      -6.2592 1.84 90  -3.394  0.0328
#Pocillopora eydouxi - Porites cylindrica             -9.0542 1.84 90  -4.910  0.0002
#Pocillopora eydouxi - Porites rus                    -9.6029 1.84 90  -5.207 <0.0001

emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols, ) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","chla_cm2_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Order and plot data by species

```{r}
#| fig.width: 10
#| fig.height: 8
#reorder based on mean values
means <- chla_avg %>% group_by(species) %>% summarize(m = mean(chla_ug_cm2_mean, na.rm = TRUE), .groups = "drop")

chla_avg_ordered <- chla_avg %>% 
  left_join(chla_summary, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

chla_plot <- ggplot() +
  geom_jitter(data = chla_avg_ordered, aes(x = full_species, y = chla_ug_cm2_mean, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = chla_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6)+
  geom_point(data = chla_summary, aes(x = full_species, y = mean), size = 2) +
  stat_summary(data = chla_avg_ordered, aes(x = full_species, y = chla_ug_cm2_mean), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "a", "a", "a", "a", "ab", "ab", "ab", "b", "b"))+
  theme_bw(base_size = 22)+
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  labs(x = "Species", color = "Species", y = expression("Chlorophyll a content" ~ (mu*g ~ cm^{-2}))) +
  ylim(0,31)
chla_plot

ggsave(here("Output", "Physiology", "chla_cm2_species_jitter.pdf"), chla_plot, h = 5, w = 10)
```

#### Chla/symbiont

Only seven species for sym data for now so edit dataframe accordinglySummarize

```{r}
chla_sym_summary <- chla_avg %>%
  group_by(full_species) %>%
  summarise(n = sum(!is.na(chla_pg_sym)),
            mean = mean(chla_pg_sym, na.rm = TRUE),
            se = se_fun(chla_pg_sym),
            .groups = "drop")
```

See if there's a difference between species

```{r}
chla.sym.spp <- lm(chla_pg_sym~full_species, data = chla_avg)
Anova(chla.sym.spp)
summary(chla.sym.spp)
check_model(chla.sym.spp)
```

Emmeans comparison and plot

```{r}
emm_obj <- emmeans::emmeans(chla.sym.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#Montipora aequituberculata - Pocillopora eydouxi    -2.94930 0.88 90  -3.352  0.0370
#Pocillopora eydouxi - Turbinaria frondens            2.95298 0.88 90   3.356  0.0365
#a, ab, ab, abc, abc, bc, c
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","chla_sym_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder and plot by species

```{r}
#| fig.width: 10
#| fig.height: 8
#reorder based on mean values
means <- chla_avg %>% group_by(species) %>% summarize(m = mean(chla_pg_sym, na.rm = TRUE), .groups = "drop")

chla_avg_ordered <- chla_avg %>% 
  left_join(chla_sym_summary, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

chla_sym_plot <- ggplot() +
  geom_jitter(data = chla_avg_ordered, aes(x = full_species, y = chla_pg_sym, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = chla_sym_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6)+
  geom_point(data = chla_sym_summary, aes(x = full_species, y = mean), size = 2) +
  #stat_summary(data = chla_avg_ordered, aes(x = full_species, y = chla_pg_sym), geom = "text", fun = max, vjust = -0.5, size = 8, label = c("a", "a", "a", "a", "a", "ab", "b"))+
  theme_bw(base_size = 22)+
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  labs(x = "Species", color = "Species", y = expression("Chlorophyll a content" ~ (pg ~ symbiont^{-2})))
  #ylim(0,8)
chla_sym_plot

ggsave(here("Output", "Physiology", "chla_sym_species_jitter.pdf"), chla_sym_plot, h = 8, w = 10)
```

## Symbiont density

Load data

```{r}
avg_sym <- read.csv(here("Data", "Physiology", "Average_Sym_Density.csv"))
# avg_sym <- avg_sym %>% 
#   filter(frag_ID != "C07") %>% 
#   filter(frag_ID != "D10") #remove crazy outliers!
```

Needed initially to calculate avg per frag sym density

```{r}
#code to calculate average Symbiont Density per fragment
#needed initially but not after writing average Symbiont Density datasheet
# phys_meta <- read.csv(here("Data", "Physiology", "Physio_meta_all.csv"))
# 
# #read in sym photo id sheet and sym counts
# counts <- read.csv(here("Data", "Physiology", "sym_counts_all.csv"))
# photos <- read.csv(here("Data", "Physiology", "sym_photos_all.csv"))
# 
# #combine
# sym_raw <- photos %>% left_join(counts, by = "image_id")
# 
# #add together counts from photos from 2 focus settings for photos
# #img_rm <- c("P8184974","P8184975","P8184976","P8174396","P8174397","P8174662","P8174663","P8174668","P8174669")
# sym_add <- sym_raw %>%
#   group_by(frag_ID,rep) %>% mutate(count_2 = sum(count)) %>%
#   filter(image_no == 1) #keep only 1 row of data for each
#   #filter(!image_id %in% img_rm) #filter out images to remove based on SE >15% of mean (as calculated below)
# 
# #average sym counts across 6 reps for each id
# #and calculate SE
# sym_avg <- sym_add %>%
#   group_by(frag_ID) %>% mutate(count_avg = mean(count_2)) %>%
#   mutate(count_se = sd(count_2)/sqrt(length(count_2))) %>%
#   #check to see which have highest error and look at add data for those >15%
#   mutate(count_var = count_se/count_avg) %>%
#   filter(rep == 6) #keep only one row of data for each
# 
# #error >0.153 = E04, H08, F07, D03
# #all error less than 0.2, just keeping all!
# #D03 - all look fine, leave
# #E04 - one count of 7, all others 12-27,
# #F07 - two counts in the 80s - remove, all others ~30
# #H08 - two counts in the 300s while all others in the 100s
# 
# #now calculate actual symbiont density
# #the 1x1mm grid should be 0.1uL
# #we want cells/mL so 10,000 x our counts (1000uL in a mL/0.1uL = x 10,000)
# #homogenate volume for all is 50mL
# #SA is calculated as well here
# #combine with metadata for calculations
# 
# sym <- sym_avg %>%
#   select(frag_ID, count_avg) %>% #keep only relevant columns
#   left_join(phys_meta, by = "frag_ID") %>% #add metadata
#   mutate(sym_mL = count_avg*10000) %>%
#   mutate(sym_total = sym_mL*blastate_vol_mL) %>%
#   mutate(sym_cm2 = sym_total/SA_cm2)
# 
# write.csv(sym, here("Data", "Physiology", "Average_Sym_Density.csv"), row.names = FALSE)
```

Summary data

```{r}
sym_summary <- avg_sym %>%
  group_by(full_species) %>%
  summarise(
    n    = sum(!is.na(sym_cm2)),
    mean = mean(sym_cm2, na.rm = TRUE),
    se   = se_fun(sym_cm2),
    .groups = "drop")

```

See if there's a difference between species

```{r}
sym.mod.spp <- lm(sym_cm2~full_species, data = avg_sym)
Anova(sym.mod.spp)
summary(sym.mod.spp)
check_model(sym.mod.spp)
```

Emmeans comparison and plot

```{r}
emm_obj <- emmeans::emmeans(sym.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
#Acropora hyacinthus - Porites cylindrica            -1261950 360000 90  -3.507  0.0235
#Echinopora lamellosa - Porites cylindrica           -1237677 360000 90  -3.439  0.0287
#Montipora vietnamensis - Porites cylindrica         -1178332 360000 90  -3.274  0.0460
#Pachyseris rugosa - Porites cylindrica              -1331170 360000 90  -3.699  0.0130
#Pocillopora eydouxi - Porites cylindrica            -1980545 360000 90  -5.504 <0.0001
#Pocillopora eydouxi - Porites rus                   -1307910 360000 90  -3.634  0.0159
#Pocillopora eydouxi - Turbinaria frondens           -1465335 360000 90  -4.072  0.0038

#a, ab, ab, ab, abc, abc, abc, bc, bc, c
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output", "Physiology","sym_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder and plot by species

```{r}
#| fig.width: 10
#| fig.height: 8
#reorder based on mean values
means <- avg_sym %>% group_by(species) %>% summarize(m = mean(sym_cm2, na.rm = TRUE), .groups = "drop")

avg_sym_ordered <- avg_sym %>% left_join(sym_summary, by = "full_species") %>% mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

sym_plot <- ggplot() +
  geom_jitter(data = avg_sym_ordered, aes(x = full_species, y = sym_cm2, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = sym_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = sym_summary, aes(x = full_species, y = mean), size = 2) +
  stat_summary(data = avg_sym_ordered, aes(x = full_species, y = sym_cm2), geom = "text", fun = max, vjust = -0.5, size = 8, label = c("a", "ab", "ab", "ab", "abc", "abc", "abc", "bc", "bc", "c"))+
  #a, ab, ab, ab, abc, abc, abc, bc, bc, c
  theme_bw(base_size = 22) +
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  ylim(0,6000000)+
  labs(x = "Species", color = "Species", y = expression("Symbiont Density" ~ (cells ~ cm^{-2})))
sym_plot

ggsave(here("Output", "Physiology", "sym_species_jitter.pdf"), sym_plot, h = 5, w = 10)
#low sym density but high chla/sym maybe helpful for P eyd here - lots of photosynthesis but lower cell counts = more energy/stores for growth
```

## Protein

Load data

```{r}
prot <- read.csv(here("Data", "Physiology", "protein_all_summary.csv"))
```

Protein summary data

```{r}
prot_summary <- prot %>%
  group_by(full_species) %>%
  summarise(
    n    = sum(!is.na(prot_ug_cm2)),
    mean = mean(prot_ug_cm2, na.rm = TRUE),
    se   = se_fun(prot_ug_cm2),
    .groups = "drop")
```

See if there's a difference between species

```{r}
prot.mod.spp <- lm(prot_ug_cm2~full_species, data = prot)
Anova(prot.mod.spp)
summary(prot.mod.spp)
check_model(prot.mod.spp)
```

Emmeans comparisons and plot

```{r}
emm_obj <- emmeans::emmeans(prot.mod.spp, ~ full_species)
emm_pairs <- pairs(emm_obj)
# Acropora hyacinthus - Turbinaria frondens            -1148.9 257 85  -4.465  0.0010
# Echinopora lamellosa - Turbinaria frondens           -1028.7 257 85  -3.998  0.0050
# Montipora aequituberculata - Turbinaria frondens     -1068.6 257 85  -4.153  0.0030
# Pachyseris rugosa - Pocillopora eydouxi                947.5 264 85   3.584  0.0190
# Pocillopora eydouxi - Porites rus                     -904.6 264 85  -3.422  0.0308
# Pocillopora eydouxi - Porites cylindrica              -867.6 264 85  -3.282  0.0457
# Pocillopora eydouxi - Turbinaria frondens            -1577.3 264 85  -5.967  <.0001
#a, ab, ab, ab, abc, abc, bc, bc, bc, c  
emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>% mutate(full_species = fct_reorder(full_species, emmean))

emm_plot <- ggplot(emm_tbl, aes(x = emmean, y = full_species, color = full_species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "none", axis.text.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols, ) + 
  labs(x = "Emmean", y = "Species")
emm_plot

#ggsave(here("Output","Physiology","prot_emmeans.pdf"), device = "pdf", height = 8, width = 8, emm_plot)
```

Reorder and plot species comparisons

```{r}
#| fig.width: 10
#| fig.height: 8
#reorder based on mean values
means <- prot %>% group_by(species) %>% summarize(m = mean(prot_ug_cm2, na.rm = TRUE), .groups = "drop")

prot_ordered <- prot %>% left_join(prot_summary, by = "full_species") %>% mutate(full_species = fct_reorder(full_species, mean))  # ascending by mean

prot_plot <- ggplot() +
  geom_jitter(data = prot_ordered, aes(x = full_species, y = prot_ug_cm2, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = prot_summary, aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = prot_summary, aes(x = full_species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  stat_summary(data = prot_ordered, aes(x = full_species, y = prot_ug_cm2), geom = "text", fun = max, vjust = -0.5, size = 8,
    label = c("a", "ab", "ab", "ab", "abc", "abc", "bc", "bc", "bc", "c"))+
  #a, ab, ab, ab, abc, abc, bc, bc, bc, c 
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  ylim(0,4100)+
  labs(x = "Species", color = "Species",
       y = expression("Protein Content" ~ (mu*g ~ cm^{-2})))
prot_plot

ggsave(here("Output", "Physiology", "prot_species_jitter.pdf"), prot_plot, h = 5, w = 10)

```

## All plots together

Put all physio plots together

```{r}
#| fig.width: 15
#| fig.height: 15
physio_plots <- ggarrange(chla_plot, chla_sym_plot, sym_plot, afdw_plot, prot_plot, nrow = 3, ncol = 2, legend = "right", common.legend = TRUE)

plot(physio_plots)
ggsave(here("Output","Physiology","physio_plots.pdf"), physio_plots, h = 15, w = 15)
```
