---
title: "SfM Analysis"
author: "Ashley Carreiro"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(vegan)
library(phyloseq)
library(car)
library(see)
library(performance)
library(emmeans)
```

Upload Data

```{r}
CoralNet <- read_csv(here("Data/SfM/CoralNetAnalysis.csv"))
CoralNetReference <- read_csv(here("Data/SfM/CoralNetGrowthForms.csv")) 
ComplexityData <- read_csv(here("Data/SfM/ComplexitySummaryStats.csv"))
TempData <- read_csv(here("Data/SfM/TempMeans.csv"))
FlowData <- read_csv(here("Data/SfM/FlowMetrics.csv"))
```


```{r}
SubstrateCoverData <- 
  left_join(CoralNet, CoralNetReference, by = c("CoralNetAbb")) %>% mutate(id = paste(SiteName, Quadrat)) %>%
    select(id, everything()) 
view(SubstrateCoverData)
```

Creating Meta Data Frame with: 
UniqueID, Site, Quadrat, complexity average metrics for each site, flow average metrics for each site, and temperature average metrics for each site

```{r}
MetaData <- SubstrateCoverData %>% 
  dplyr::select(id, SiteName, Year, Quadrat) %>%
  group_by(id) %>%
  summarise(across(everything(), ~ paste(unique(.), collapse = ", ")), .groups = "drop") %>% #Collapse all the rows that have the same data across all columns
  left_join(ComplexityData) %>%
  left_join(FlowData) %>%
  left_join(TempData) %>%
  
  
  mutate(TempRange = maxTemp - minTemp) %>%
  mutate(FlowRange = MaxSpeed - MinSpeed)
            
            
```


```{r}
SubstrateCoverDataWide <- SubstrateCoverData %>%

  mutate(Count = 1) %>%
    group_by(id, CoralNetAbb) %>%
  select(id,CoralNetAbb, Count) %>%
  pivot_wider(names_from = CoralNetAbb, values_from = Count,
            values_fill = 0,
            values_fn = sum) %>%
  mutate(Total = sum(c_across(TurfAl:Zoanthids), na.rm = TRUE)) 
  

SubstrateCoverDataWide$Total #making sure there is 100 
```

Function to Creat Dataframes of everythign we want to graph, 
DOnʻt think this was necessary can just use SbustrateCoverMetaData

```{r}
PercentCover_Fun <- function(  G, rowx, rowy, GroupName) 
{
  
SubstrateCoverData %>%
    mutate(Count = 1) %>%
    group_by(id, {{G}}) %>%
    select(id, {{G}}, Count) %>%
    pivot_wider(names_from = {{G}}, values_from = Count,
                values_fill = 0,
                values_fn = sum) %>%
    mutate(Total = sum(c_across({{rowx}}:{{rowy}}), na.rm = TRUE)) %>%
    mutate(across({{rowx}}:{{rowy}}, ~ (.x / Total))) %>%
    pivot_longer(cols = {{rowx}}:{{rowy}}, 
                 names_to = GroupName) %>%
    left_join(MetaData, by = join_by(id)) %>%
    filter(value != 0)
}

PercentCoverMeta <- PercentCover_Fun( CoralNetAbb, TurfAl, Zoanthids, GroupName = "CoralNetAbb")

#FunctionalGroupMeta <- PercentCover_Fun( FunctionalGroup, Algae, Seagrass, GroupName = "FunctionalGroup")




#GrowthFormMeta <- PercentCover_Fun(GrowthForm, CTB, Seagrass, GroupName = "GrowthForm") %>% 




HardCoralMeta <- PercentCoverMeta %>%
  filter(FunctionalGroup == "Hard coral") 
  


```
  
Visualize the Data, create a funciton to look at all the different meta data variables to see if there is any relationship 

```{r}
RelationshipPlots_fun <- function(Df, MetaVariable, PlotName)
  
{

  ggplot(Df, aes(x= {{MetaVariable}}, y = value, color = SiteName)) +
         geom_point() +
     labs( y = "PercentCover") +
    
  
  ggsave(here("Output", "SfM", PlotName))
}

RelationshipPlots_fun(HardCoralMeta, MeanRug, PlotName = "HardCoralRug.png")
RelationshipPlots_fun(HardCoralMeta, MeanHR, PlotName = "HardCoralHeightRange.png")
RelationshipPlots_fun(HardCoralMeta, MeanDepth, PlotName = "HardCoralMeanDepth.png")
RelationshipPlots_fun(HardCoralMeta, TempRange, PlotName = "HardCoralTempRange.png")
RelationshipPlots_fun(HardCoralMeta, MeanSpeed, PlotName = "HardCoralMeanSpeed.png")
RelationshipPlots_fun(HardCoralMeta, FlowRange, PlotName = "HardCoralFlowRange.png")
RelationshipPlots_fun(HardCoralMeta, MaxSpeed, PlotName = "HardCoralMaxSpeed.png")


```

```{r}
RelationshipBarPlots_fun <- function(Df, MetaVariable, Variables)
  
{

  ggplot(Df, aes( x= {{MetaVariable}}, y = value, fill = {{Variables}})) +
         geom_col(position = "stack") +
    labs( y = "PercentCover") 
}


RelationshipBarPlots_fun(PercentCoverMeta, MeanRug, FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, MeanHR, FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, MeanDepth, FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, TempRange , FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, MeanSpeed, FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, FlowRange, FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, MaxSpeed, FunctionalGroup)

RelationshipBarPlots_fun(HardCoralMeta, MeanRug, GrowthForm)
RelationshipBarPlots_fun(HardCoralMeta, MeanHR, GrowthForm)
RelationshipBarPlots_fun(HardCoralMeta, MeanDepth, GrowthForm)
RelationshipBarPlots_fun(HardCoralMeta, TempRange , GrowthForm)
RelationshipBarPlots_fun(HardCoralMeta, MeanSpeed, GrowthForm)
RelationshipBarPlots_fun(HardCoralMeta, FlowRange, GrowthForm)
RelationshipBarPlots_fun(HardCoralMeta, MaxSpeed, GrowthForm)
```







```{r}
count_data <- SubstrateCoverDataWide %>% dplyr::select(-Total) %>% mutate(across(c(TurfAl:Zoanthids),as.numeric))
rownames(count_data) <- count_data$id
newcountdata <- count_data[,-1]

#has to match column names with variable
taxa <- CoralNetReference
rownames(taxa) <- taxa$CoralNetAbb

#has to match id with count data
samdf <- MetaData 
rownames(samdf) <- samdf$id #Adding rownames 

#BiocManager::install("phyloseq")

mcounts <- as.matrix(newcountdata)
mtaxa <- as.matrix(taxa)

ps <- phyloseq(otu_table(mcounts, taxa_are_rows = F), #Making phyloseq  object
               sample_data(samdf),
               tax_table(mtaxa))


ps.rel <- transform_sample_counts(ps, function(x) x /sum(x)) #Making relative abundance dataframe, trasnforming to divide by the sum to make it relative abundance, for each sample (combined quadrat and site)

##DO THIS AT ALL DIFFERENT LEVELS###

#make different plots with functional group, species
plot_bar(ps.rel, x = "id", y = "Abundance", fill = "FunctionalGroup")+
  facet_wrap(~SiteName, scales = "free") +
  geom_bar(stat = "identity")

#filter dataframe to just look at coral
ps.coral <- subset_taxa(ps, FunctionalGroup == "Hard coral")
ps.rel.coral <- subset_taxa(ps.rel, FunctionalGroup == "Hard coral")

SubstrateCoverFullDataFrame <- psmelt(ps.rel) %>%
  filter(Abundance != 0)

ps.rel.cover <- merge_taxa(ps.rel, "FunctionalGroup")
FunctionalGroupFullDataFrame <- psmelt(ps.rel.cover) %>%
  filter(Abundance != 0)


ggplot(SubstrateCoverFullDataFrame aes(x = MeanRug, y = ))

plot_bar(ps.rel.coral, x = "id", y = "Abundance", fill = "Name")+
  facet_wrap(~SiteName, scales = "free") +
  geom_bar(stat = "identity")
  
#plot at dif taxonomic level
ps.rel.coral_growth <- tax_glom(ps.rel.coral, "GrowthForm")
coralgrowth_melt <- psmelt(ps.rel.coral_growth)
plot_bar(ps.rel.coral_growth, x = "id", y = "Abundance", fill = "GrowthForm")+
  facet_wrap(~SiteName, scales = "free") +
  geom_bar(stat = "identity")
```









```{r}
PercentCoverMeta <- PercentCoverData %>%
  pivot_longer(cols = TurfAl:Zoanthids, 
               names_to = "CoralNetAbb") %>%
  left_join(MetaData, by = join_by(id)) %>%
  left_join(CoralNetReference)
```
```{r}
rel_abun <- as.data.frame(ps.rel@otu_table)
rel_abun$sample_id <- rownames(rel_abun)
RelativeAbundance <- left_join(DiversityMetrics, rel_abun)
```

```{r}
ggplot(PercentCoverLong, aes(x= SiteName, y = value, fill = CoralNetAbb)) +
         geom_bar(stat=("")) 
```

```{r}
ggplot(TotalSitePercentCover, aes(x= SiteName, y = value, fill = CoralNetAbb)) +
         geom_bar(stat="sum") 
```

DIVERSITY

```{r}
CoralDiversity <- data.frame(estimate_richness(ps.coral, split=TRUE, measures=c("Shannon","Observed")))

CoralDiversity$sample_id <- rownames(CoralDiversity)

samdf <- data.frame(sample_data(ps.coral))
samdf$sample_id <- rownames(samdf)
Diversity <- merge(CoralDiversity,samdf, by = "sample_id") #add sample data

#add evenness = shannon diversity divided by species richness
Diversity$even <- Diversity$Shannon/(log(Diversity$Observed))

#SAVE DATAFRAME
write.csv(Diversity, file = here("Data/SfM/CoralNetDiversityMetrics.csv"), row.names = FALSE)


```

Read in the new dataframe

```{r}
DiversityMetrics <- read.csv(here("Data/SfM/CoralNetDiversityMetrics.csv"))
```


```{r}
#Shannon by species across everything
gg.spp.sha <- ggplot(DiversityMetrics,aes(x=SiteName,y=Shannon,fill=SiteName))+
  geom_boxplot(outlier.shape=NA,size=1)+
  theme_classic(base_size=22)+
  geom_jitter(alpha=0.5)+
  #scale_fill_manual("Coral Species", values = c("darkkhaki","darksalmon","indianred"), labels = c(substitute(paste(italic("Porites sp."))), substitute(paste(italic("S. radians"))),substitute(paste(italic("S. siderea")))))+
 # ylab("Shannon Index")+
  #xlab("Coral Species")+
  theme(legend.position="right",axis.text.x = element_blank(),axis.ticks.x=element_blank())
  #facet_wrap(m_y~.,scales = "free_x")
gg.spp.sha
```
```{r}
#Shannon by species across everything
gg.spp.rich <- ggplot(DiversityMetrics,aes(x=SiteName,y=Observed,fill=SiteName))+
  geom_boxplot(outlier.shape=NA,size=1)+
  theme_classic(base_size=22)+
  geom_jitter(alpha=0.5)+
  #scale_fill_manual("Coral Species", values = c("darkkhaki","darksalmon","indianred"), labels = c(substitute(paste(italic("Porites sp."))), substitute(paste(italic("S. radians"))),substitute(paste(italic("S. siderea")))))+
 # ylab("Shannon Index")+
  #xlab("Coral Species")+
  theme(legend.position="right",axis.text.x = element_blank(),axis.ticks.x=element_blank())
  #facet_wrap(m_y~.,scales = "free_x")
gg.spp.rich
```

Make Regression Plot 
#DO THIS FOR EVERYTHING TO SEE IF THERE IS ANYTYPE OF PATTERN

```{r}
ggplot(DiversityMetrics, aes(x = MeanRug, y = Observed, color = SiteName)) +
  geom_point()
```




```{r}
#Shannon by species across everything
gg.spp.Evenness <- ggplot(DiversityMetrics,aes(x=SiteName,y=even,fill=SiteName))+
  geom_boxplot(outlier.shape=NA,size=1)+
  theme_classic(base_size=22)+
  geom_jitter(alpha=0.5)+
  #scale_fill_manual("Coral Species", values = c("darkkhaki","darksalmon","indianred"), labels = c(substitute(paste(italic("Porites sp."))), substitute(paste(italic("S. radians"))),substitute(paste(italic("S. siderea")))))+
 # ylab("Shannon Index")+
  #xlab("Coral Species")+
  theme(legend.position="right",axis.text.x = element_blank(),axis.ticks.x=element_blank())
  #facet_wrap(m_y~.,scales = "free_x")
gg.spp.Evenness

ggsave(here("Output", "SfM", "DiversityPlot_Evenness.png"),
         width = 6, height = 6)
```
Linear model looking across site

Oki04 is different from all 

```{r}
shannonModel <- lm(Shannon ~ SiteName, data = DiversityMetrics)
anova(shannonModel)
summary(shannonModel)
check_model(shannonModel)

ShannonEmm <- emmeans::emmeans(shannonModel, ~SiteName)
emm_pairs <- pairs(ShannonEmm)
view(emm_pairs)

write.csv(emm_pairs, here("Output/SfM/ShannonDiversityStats.csv"))

```

```{r}

EvennessModel <- lm(even ~ SiteName, data = DiversityMetrics)
anova(EvennessModel)
summary(EvennessModel)
check_model(EvennessModel)

EvennessEmm <- emmeans::emmeans(EvennessModel, ~SiteName)
Evennessemm_pairs <- pairs(EvennessEmm)
view(Evennessemm_pairs)

write.csv(Evennessemm_pairs, here("Output/SfM/EvennessDiversityStats.csv"))

```

Organize the Complexity  data 
First read in the data that was created in the other script


BETA DIVERSITY




