---
title: "SfM Analysis"
author: "Ashley Carreiro"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(vegan)
library(phyloseq)
library(car)
library(see)
library(performance)
library(emmeans)
```

Upload Data

```{r}
CoralNet <- read_csv(here("Data/SfM/CoralNetAnalysis.csv"))
CoralNetReference <- read_csv(here("Data/SfM/CoralNetGrowthForms.csv")) 
ComplexityData <- read_csv(here("Data/SfM/ComplexitySummaryStats.csv"))
TempData <- read_csv(here("Data/SfM/TempMeans.csv"))
FlowData <- read_csv(here("Data/SfM/FlowMetrics.csv"))
```


```{r}
SubstrateCoverData <- 
  left_join(CoralNet, CoralNetReference, by = c("CoralNetAbb")) %>% mutate(id = paste(SiteName, Quadrat)) %>%
    select(id, everything()) 
view(SubstrateCoverData)
```

Creating Meta Data Frame with: 
UniqueID, Site, Quadrat, complexity average metrics for each site, flow average metrics for each site, and temperature average metrics for each site

```{r}
MetaData <- SubstrateCoverData %>% 
  dplyr::select(id, SiteName, Year, Quadrat) %>%
  group_by(id) %>%
  summarise(across(everything(), ~ paste(unique(.), collapse = ", ")), .groups = "drop") %>% #Collapse all the rows that have the same data across all columns
  left_join(ComplexityData) %>%
  left_join(FlowData) %>%
  left_join(TempData) %>%
  
  
  mutate(TempRange = maxTemp - minTemp) %>%
  mutate(FlowRange = MaxSpeed - MinSpeed)
            
SiteMetaData <- SubstrateCoverData %>% 
  dplyr::select(SiteName) %>%
  group_by(SiteName) %>%
  summarise(across(everything(), ~ paste(unique(.), collapse = ", ")), .groups = "drop") %>% #Collapse all the rows that have the same data across all columns
  left_join(ComplexityData) %>%
  left_join(FlowData) %>%
  left_join(TempData) %>%
  
  
  mutate(TempRange = maxTemp - minTemp) %>%
  mutate(FlowRange = MaxSpeed - MinSpeed)            
```


Function to Creat Dataframes of everythign we want to graph, 


```{r}
PercentCover_Fun <- function( G, rowx, rowy, GroupName) 
{
  
SubstrateCoverData %>%
    mutate(Count = 1) %>%
    group_by(id, {{G}}) %>%
    select(id, {{G}}, Count) %>%
    pivot_wider(names_from = {{G}}, values_from = Count,
                values_fill = 0,
                values_fn = sum) %>%
    mutate(Total = sum(c_across({{rowx}}:{{rowy}}), na.rm = TRUE)) %>%
    mutate(across({{rowx}}:{{rowy}}, ~ (.x / Total))) %>%
    pivot_longer(cols = {{rowx}}:{{rowy}}, 
                 names_to = GroupName) %>%
    left_join(MetaData, by = join_by(id)) %>%
    filter(value != 0)
}

QuadratePercentCoverMeta <- PercentCover_Fun( CoralNetAbb, TurfAl, Zoanthids, GroupName = "CoralNetAbb")

SiteFunctionalPercentCover <- PercentCover_Fun(SiteName, FunctionalGroup, Algae, Seagrass, GroupName = "FunctionalGroup", SiteMetaData)



HardCoralMeta <- PercentCoverMeta %>%
  filter(FunctionalGroup == "Hard coral") 
  
SiteCover_Fun <- function(  G, rowx, rowy, GroupName) 
{
  
SubstrateCoverData %>%
    mutate(Count = 1) %>%
    group_by(SiteName, {{G}}) %>%
    select(SiteName, {{G}}, Count) %>%
    pivot_wider(names_from = {{G}}, values_from = Count,
                values_fill = 0,
                values_fn = sum) %>%
    mutate(Total = sum(c_across({{rowx}}:{{rowy}}), na.rm = TRUE)) %>%
    mutate(across({{rowx}}:{{rowy}}, ~ (.x / Total))) %>%
    pivot_longer(cols = {{rowx}}:{{rowy}}, 
                 names_to = GroupName) %>%
    left_join(SiteMetaData, by = join_by(SiteName)) %>%
    filter(value != 0)
}

TotalSubstrateSiteMeta <- SiteCover_Fun(CoralNetAbb, TurfAl, Zoanthids, 
                                        GroupName = "CoralNetAbb")
  
FunctionalSiteMeta <- SiteCover_Fun( FunctionalGroup, Algae, Seagrass, GroupName = "FunctionalGroup")

HCFunctionalSiteMeta <- FunctionalSiteMeta %>%
  filter(FunctionalGroup == "Hard coral")

GrowthFormSiteMeta <- SiteCover_Fun(GrowthForm, CTB, Seagrass, 
                                    GroupName = "GrowthForm")

HCGrowthFormSite <- GrowthFormSiteMeta %>%
  filter(!GrowthForm %in% c("CTB", "Clam", "Disease", "CCA", "Sponge", "NA", "RDC", "Algae", "Sand", "Zooanthid", "Sea Cucumber", "Millepora", "Ascidian", "Seagrass"))




```
  
Visualize the Data, create a funciton to look at all the different meta data variables to see if there is any relationship 

```{r}
RelationshipPlots_fun <- function(Df, MetaVariable, PlotName)
  
{

  ggplot(Df, aes(x= {{MetaVariable}}, y = value, color = SiteName)) +
         geom_point(size = 3) +
     labs( y = "HardCoralPercentCover") +
    
  
  ggsave(here("Output", "SfM", PlotName))
}

RelationshipPlots_fun(HCFunctionalSiteMeta, MeanRug, PlotName = "HardCoralRug.png")
RelationshipPlots_fun(HCFunctionalSiteMeta, MeanHR, PlotName = "HardCoralHeightRange.png")
RelationshipPlots_fun(HCFunctionalSiteMeta, MeanDepth, PlotName = "HardCoralMeanDepth.png")
RelationshipPlots_fun(HCFunctionalSiteMeta, TempRange, PlotName = "HardCoralTempRange.png")
RelationshipPlots_fun(HCFunctionalSiteMeta, MeanSpeed, PlotName = "HardCoralMeanSpeed.png")
RelationshipPlots_fun(HCFunctionalSiteMeta, FlowRange, PlotName = "HardCoralFlowRange.png")
RelationshipPlots_fun(HCFunctionalSiteMeta, MaxSpeed, PlotName = "HardCoralMaxSpeed.png")


```

```{r}
RelationshipBarPlots_fun <- function(Df, MetaVariable, Variables)
  
{
  ggplot(Df, aes(x = {{MetaVariable}}, y = value, fill = {{Variables}})) +
    geom_col(position = "stack") +
    
    labs(y = "PercentCover")


}

          
          
RelationshipBarPlots_fun(TotalSubstrateSiteMeta, MeanRug, CoralNetAbb)
RelationshipBarPlots_fun(TotalSubstrateSiteMeta, MeanHR, CoralNetAbb)
RelationshipBarPlots_fun(PercentCoverMeta, MeanDepth, FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, TempRange , FunctionalGroup)
RelationshipBarPlots_fun(PercentCoverMeta, MeanSpeed, FunctionalGroup)
RelationshipBarPlots_fun(TotalSubstrateSiteMeta, FlowRange, CoralNetAbb)
RelationshipBarPlots_fun(PercentCoverMeta, MaxSpeed, FunctionalGroup)

RelationshipBarPlots_fun(HCGrowthFormSite, SiteName, GrowthForm)

RelationshipBarPlots_fun(HCGrowthFormSite, MeanRug, GrowthForm)
RelationshipBarPlots_fun(HCGrowthFormSite, MeanHR, GrowthForm)
RelationshipBarPlots_fun(HCGrowthFormSite, MeanDepth, GrowthForm)
RelationshipBarPlots_fun(HCGrowthFormSite, TempRange , GrowthForm)
RelationshipBarPlots_fun(HCGrowthFormSite, MeanSpeed, GrowthForm)
RelationshipBarPlots_fun(HCGrowthFormSite, FlowRange, GrowthForm)
RelationshipBarPlots_fun(HCGrowthFormSite, MaxSpeed, GrowthForm)
```





