---
title: "Tilt Meter Analysis"
author: "Ashley Carreiro"
format: html
editor: visual
---

# Load Libraries

```{r}
library(tidyverse)
library(here)
library(climaemet)
library(patchwork)
library(ggplotify)
library(magick)
library(grid)

```

## Ignore: Creating New CSV Files

Trimming down the tilt meter data to one hour after deployment and one hour before retrieval of the tilt meters.

Making sure all data is captured at one minute intervals.

Note that all data has been adjusted for the Magnetic Declination of -5.85 West within the Domino app.

All data speed units are in cm/s

**This code does not need to be run again!**

```{r}
#| eval: false
#| echo: false
read_csv(here("Data/TiltMeterData/Tilt1_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% #Changing the headers int he dataframe
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% #Changing the date to character using lubridatde
  filter(between(DateTime, ymd_hms("2025-08-01 11:21:13"), ymd_hms("2025-08-14 08:45:13"))) %>% #Filtering between the dates that we want data from
  write_csv(file ="Data/TiltMeterData/Tilt1_Current.csv") #Creating new CSV File with filtered dates

read_csv(here("Data/TiltMeterData/Tillt2_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 13:00:42"), ymd_hms("2025-08-14 09:30:42"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt2_Current.csv")

read_csv(here("Data/TiltMeterData/Tilt3_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 14:45:50"), ymd_hms("2025-08-14T11:30:50"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt3_Current.csv")

#Tilt 4 was taking data every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt4_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 10:45:14"), ymd_hms("2025-08-14 12:45:04"))) %>% 
    group_by(DateTime = floor_date(DateTime, "min")) %>% #Averaging the data over 1 minute to match the other data
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(file ="Data/TiltMeterData/Tilt4_Current.csv")  

read_csv(here("Data/TiltMeterData/Tilt5_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 12:18:28"), ymd_hms("2025-08-15 12:30:28"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt5_Current.csv")

read_csv(here("Data/TiltMeterData/Tilt6_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 13:57:00"), ymd_hms("2025-08-15 11:30:00"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt6_Current.csv")

#Tilt 7 was taking readings every second
read_csv(here("Data/TiltMeterData/Tilt7_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 10:32:32"), ymd_hms("2025-08-15 08:46:00"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(file ="Data/TiltMeterData/Tilt7_Current.csv")

#Tilt 8 was taking readings every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt8_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 12:06:06"), ymd_hms("2025-08-15 9:47:06"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(file ="Data/TiltMeterData/Tilt8_Current.csv")
```

Temperature Data

```{r}
#| eval: false
#| echo: false
 read_csv(here("Data/TiltMeterData/Tlt1_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% #Changing the headers int he dataframe
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% #Changing the date to character using lubridatde
  filter(between(DateTime, ymd_hms("2025-08-01 11:21:13"), ymd_hms("2025-08-14 08:45:13"))) %>%  #Filtering between the dates that we want data from
  write_csv(here("Data/TiltMeterData/Tilt1_Temperature.csv")) #Creating new CSV File with filtered dates

 read_csv(here("Data/TiltMeterData/Tilt2_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 13:00:42"), ymd_hms("2025-08-14 09:30:42"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt2_Temperature.csv"))
 
 read_csv(here("Data/TiltMeterData/Tilt3_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 14:45:50"), ymd_hms("2025-08-14T11:30:50"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt3_Temperature.csv"))
 
 #Tilt 4 was taking data every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt4_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 10:45:14"), ymd_hms("2025-08-14 12:45:04"))) %>% 
    group_by(DateTime = floor_date(DateTime, "min")) %>% #Averaging the data over 1 minute to match the other data
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(here("Data/TiltMeterData/Tilt4_Temperature.csv"))
 
read_csv(here("Data/TiltMeterData/Tilt5_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 12:18:28"), ymd_hms("2025-08-15 12:30:28"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt5_Temperature.csv"))

read_csv(here("Data/TiltMeterData/Tilt6_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 13:57:00"), ymd_hms("2025-08-15 11:30:00"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt6_Temperature.csv"))

#Tilt 7 was taking readings every second
read_csv(here("Data/TiltMeterData/Tilt7_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 10:32:32"), ymd_hms("2025-08-15 08:46:00"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(here("Data/TiltMeterData/Tilt7_Temperature.csv"))

#Tilt 8 was taking readings every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt8_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 12:06:06"), ymd_hms("2025-08-15 9:47:06"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(here("Data/TiltMeterData/Tilt8_Temperature.csv"))

```

Doing the same for the tide data

Tide data was downloaded from: <https://www.gsi.go.jp/kanshi/tide_data_21_e.html>

```{r}
#| eval: false
#| echo: false
read_table(here("Data/TiltMeterData/21_2021-2025_hour.txt"), col_names = FALSE) %>%
select(-1) %>% #Drops the first column
set_names(c("Date", "Hour", "Sealevel_mm", "DatumConstant_mm", "FixedPointHeight_mm")) %>%
 mutate(Date = lubridate::ymd(Date)) %>% 
  filter(between(Date, ymd("2025-08-01"), ymd("2025-08-15"))) %>%
  
write_csv(here("Data/TiltMeterData/TideData_2025.08.csv"))

```

## Load Data

Load current, temperature, and tide data from new csv files created above.

```{r}
#Load Current Data
Tilt1Data <- read_csv(here("Data", "TiltMeterData", "Tilt1_Current.csv"))
Tilt2Data <- read_csv(here("Data", "TiltMeterData", "Tilt2_Current.csv"))
Tilt3Data <- read_csv(here("Data","TiltMeterData", "Tilt3_Current.csv"))
Tilt4Data <- read_csv(here("Data","TiltMeterData","Tilt4_Current.csv"))
Tilt5Data <- read_csv(here("Data","TiltMeterData", "Tilt5_Current.csv"))
Tilt6Data <- read_csv(here("Data","TiltMeterData", "Tilt6_Current.csv"))
Tilt7Data <- read_csv(here("Data","TiltMeterData", "Tilt7_Current.csv"))
Tilt8Data <- read_csv(here("Data","TiltMeterData", "Tilt8_Current.csv"))

#Load Temp Data

Tilt1Temp <- read_csv(here("Data/TiltMeterData/Tilt1_Temperature.csv"))
Tilt2Temp <- read_csv(here("Data/TiltMeterData/Tilt2_Temperature.csv"))
Tilt3Temp <- read_csv(here("Data/TiltMeterData/Tilt3_Temperature.csv"))
Tilt4Temp <- read_csv(here("Data/TiltMeterData/Tilt4_Temperature.csv"))
Tilt5Temp <- read_csv(here("Data/TiltMeterData/Tilt5_Temperature.csv"))
Tilt6Temp <- read_csv(here("Data/TiltMeterData/Tilt6_Temperature.csv"))
Tilt7Temp <- read_csv(here("Data/TiltMeterData/Tilt7_Temperature.csv"))
Tilt8Temp <- read_csv(here("Data/TiltMeterData/Tilt8_Temperature.csv"))

#Load Tide Data

TideData <- read_csv(here("Data/TiltMeterData/TideData_2025.08.csv"))

#Read in Pictures

Tilt1Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt1jpg.jpg"))
Tilt2Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt2.jpg"))
Tilt3Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt3.jpg"))
Tilt4Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt4.jpg"))
Tilt5Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt5.jpg"))
Tilt6Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt6.jpg"))
Tilt7Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt7.jpg"))
Tilt8Photo <- image_read(here("Output/TiltMeterData/TiltPhotos/Tilt8.jpg"))



```

# Organizing the data

### Tilt Meter Data

Creating a function to organize all the current data

-   Separating Data

-   

```{r}
Tilt_Data_fun <- function(df, tilt_name){  #Function to add a tilt name column and separate data and time column 
df %>%
    
separate(col = DateTime, #Choose the column that you want separated
             into = c("Date", "Time"), # Separate it into two columns "Data" and "Time"
             sep = " ", #What are the two separated by 
             remove = FALSE) %>% #Keep the orginial DateTime Column    
    
mutate( #Set datetime,a dding in the UTC timezone so wer are removing taht
       Date = lubridate::ymd(Date), #Set Date
       Time = lubridate::hms(Time), #Set Time
       Speed_m = Speed * 0.01, #Change cm/s -> m/s
       VelocityN_m = Velocity_N * 0.01, #Change cm/s -> m/s
       VelocityE_m = Velocity_E* 0.01, #Change cm/s -> m/s
   
       #Double check that there are no negative headings, can do 360 + negative
       
       TiltMeter = tilt_name) %>%
    
    select(TiltMeter, everything()) #Moving TiltMeter# Column to be the first column
    
}

Tilt1DataClean <- Tilt_Data_fun(Tilt1Data, "Tilt1")
Tilt2DataClean <- Tilt_Data_fun(Tilt2Data, "Tilt2")
Tilt3DataClean <- Tilt_Data_fun(Tilt3Data, "Tilt3")
Tilt4DataClean <- Tilt_Data_fun(Tilt4Data, "Tilt4")
Tilt5DataClean <- Tilt_Data_fun(Tilt5Data, "Tilt5")
Tilt6DataClean <- Tilt_Data_fun(Tilt6Data, "Tilt6")
Tilt7DataClean <- Tilt_Data_fun(Tilt7Data, "Tilt7")
Tilt8DataClean <- Tilt_Data_fun(Tilt8Data, "Tilt8")
```

#### Summary Stats 

First, join all the Current Data

```{r}
JoinedData <- rbind(Tilt1DataClean, Tilt2DataClean, Tilt3DataClean, Tilt4DataClean, Tilt5DataClean,  Tilt6DataClean,  Tilt7DataClean, Tilt8DataClean)

```

Visualize Summary stats to see where to filter data from **Don't Need to Run Everytime**

```{r}
summary(JoinedData$Speed)
lower_bound <- quantile(JoinedData$Speed, 0.001)
Upper_bound <- quantile(JoinedData$Speed, 0.999)
Upper_bound#Upper bound is 25.01416 we will filter data from this number 

boxplot.stats(JoinedData$Speed_m)

sum(JoinedData$Speed >= .5) #Counting the number of points above .5
# = 5

sum(JoinedData$Speed> 25.01416) #Number of points above the upper bound, 0.25, 145 points

table(cut(JoinedData$Speed_m , seq(0, .25,.05), include.lowest = TRUE)) ##Showing me the different bins and how many data points I have in each one

Means <- JoinedData %>%
  group_by(TiltMeter) %>%
  filter(Speed <=  25.01416) %>%
  summarise(mean(Speed))
Means #Giving the average speed m/s of each tilt meter
max(JoinedData$Speed)
#Visualize the data 

ggplot(JoinedData, aes(x = DateTime, y = Speed_m)) +
  geom_point() +
   facet_wrap(~TiltMeter)
```

Now we will average everything by hour and create a new data frame **hourly**

This dataframe is in cm/s

Filtering out the 99.9% upper bound which is anything above 25.01416 cm/s

```{r}
HourlyTilt_Data_fun <- function(df){  #Function to add a tilt name column and separate data and time column 
 df %>% 
    
    group_by(DateTime = floor_date(DateTime, "hour")) %>% #New datetime that is rounded to the start of each hour
  summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
    filter(Speed <=  25.01416) 
}

Tilt1DataHourly <- HourlyTilt_Data_fun(Tilt1Data)
Tilt2DataHourly <- HourlyTilt_Data_fun(Tilt2Data)
Tilt3DataHourly <- HourlyTilt_Data_fun(Tilt3Data)
Tilt4DataHourly <- HourlyTilt_Data_fun(Tilt4Data)
Tilt5DataHourly <- HourlyTilt_Data_fun(Tilt5Data)
Tilt6DataHourly <- HourlyTilt_Data_fun(Tilt6Data)
Tilt7DataHourly <- HourlyTilt_Data_fun(Tilt7Data)
Tilt8DataHourly <- HourlyTilt_Data_fun(Tilt8Data)

```

### Temperature Data

```{r}
Tilt_Temp_fun <- function(df){ 
df %>%
    
separate(col = DateTime, #Choose the column that you want separated
         into = c("Date", "Time"), # Separate it into two columns "Data" and "Time"
         sep = " ", #What are the two separated by 
         remove = FALSE) %>% #Keep the orginial DateTime Column    
  mutate(DateTime = lubridate::ymd_hms(DateTime), #Set datetime
         Date = lubridate::ymd(Date), #Set Date
         Time = lubridate::hms(Time))
}

Tilt1TempData <- Tilt_Temp_fun(Tilt1Temp)
Tilt2TempData <- Tilt_Temp_fun(Tilt2Temp)
Tilt3TempData <- Tilt_Temp_fun(Tilt3Temp)
Tilt4TempData <- Tilt_Temp_fun(Tilt4Temp)
Tilt5TempData <- Tilt_Temp_fun(Tilt5Temp)
Tilt6TempData <- Tilt_Temp_fun(Tilt6Temp)
Tilt7TempData <- Tilt_Temp_fun(Tilt7Temp)
Tilt8TempData <- Tilt_Temp_fun(Tilt8Temp)

bad_dates <- Tilt7TempData %>%
  filter(is.na(lubridate::ymd_hms(DateTime)))
bad_dates2 <- Tilt7DataClean %>%
  filter(is.na(lubridate::ymd_hms(DateTime)))
```

### Tide Data

```{r}
 TideDataHourly <- TideData %>%
   mutate(DateTime = ymd_hms(paste(Date, Hour, "00:00")), #Combining Data and time into one column 
          Sealevel_m = Sealevel_mm / 1000) %>%
    select(DateTime, everything())

```

# Plotting the Data

We are going to make a big function to make a plot of the following

-   Feather Plot. \| Wind Rose Plot

-   Temperature Plot \| Elipses Plot?

-   Tide Plot \| Photo of Tilt Meter?

```{r}
Velocity_TidePlot_fun <- function(Tiltdf, FeatherPlotname, Tempdf, TempPlotname, Tiltdf2, Sitename1, AverageSpeed, WindrosePlotname, TiltPhoto, combinedplot, Sitename) {
  
time_limits <- range(Tiltdf$DateTime, na.rm = TRUE)

FeatherPlot <- ggplot(Tiltdf, aes( DatetTime, y = 0)) +
 
  geom_segment(aes( x = DateTime, y = 0, 
                    xend = DateTime, 
                    yend = Velocity_N, #Makes the arrows point up/down
                    color = "Velocity North"), 
               
               lineend = "round" ,
               linewidth = 1,
               alpha = 0.5) +
  
  geom_segment(aes( x = DateTime, y = 0,
                    xend = DateTime, #Makes the arrows lean left/right
                    yend = Velocity_E, #Makes the arrows point up/down
                    color = "Velocity East"), 
               #arrow = arrow(length = unit(0.2, "cm")),
               linewidth = 1, 
               alpha = 0.25) +
  geom_point(aes(x = DateTime, y = Velocity_N), color = "blue4", size = 0.5) +
  geom_point(aes(x = DateTime, y = Velocity_E), color = "coral2", size = 0.5) +
  
  scale_color_manual(name = "Velocity Direction", 
                     values = c("Velocity North" = "blue4", "Velocity East" = "coral2")) +
  
   scale_x_datetime(limits = time_limits, 
                   date_breaks = "2 day", 
                   minor_breaks = "1 day", 
                   date_labels = ("%b %d")) +
 
  labs( x = "Date",
    y = "Velocity (cm/s)") +
  theme_minimal() +
   theme( legend.position = "inside", #axis.text.x = element_text(angle = 45, hjust =1) +
        legend.position.inside = c(0.5, 0.0), 
        legend.box.background = element_rect(color = "black"), 
        #legend.key.size = unit(0.1, "cm"), 
           axis.text.x = element_text(angle = 45, hjust =1))
       
 
 ggsave(here("Output","TiltMeterData",FeatherPlotname))
 
 
 TempPlot <- ggplot(Tempdf, aes(DateTime, Temperature)) +
  geom_line(color = "pink3") +
    scale_x_datetime(limits = time_limits, 
                   date_breaks = "2 day", 
                   minor_breaks = "1 day", 
                   date_labels = ("%b %d")) +
     labs(y = "Temperature (°C)") + 
   
     theme_minimal() +
   theme( axis.title.x = element_blank(), 
          axis.text.x = element_blank())
          
     
  
ggsave(here("Output", "TiltMeterData", TempPlotname))
 


TidePlot <- ggplot(TideDataHourly, aes(DateTime, Sealevel_m)) +
   geom_line(color = "steelblue") +
   scale_x_datetime(limits = time_limits, 
                   date_breaks = "2 day", 
                   minor_breaks = "1 day", 
                   date_labels = ("%b %d")) +
  labs( y = "Sealevel (m) ", 
        x = " Date") +
  
  
       theme_minimal() +
      theme( axis.text.x = element_text(angle = 45, hjust =1))
  
   
  
 
 WindrosePlot <- ggwindrose(
    speed = Tiltdf2$Speed, 
    direction = Tiltdf2$Heading,
    n_directions = 8,
    speed_cuts = seq(0, 25, 5), Inf) +
    #stack_revers = TRUE) + #If we want the higher speeds on the outside of the plot
  

    scale_fill_manual(
      values = c( "skyblue","orange" ,"green2", "blue","red", "darkred"),
      drop = FALSE
    ) +

   
      theme_minimal() +
   labs(subtitle = AverageSpeed, fill = "Current Speed\n (cm/s)", 
        title = Sitename1) +
    theme(
      axis.title =  element_blank(),
      text = element_text(face = "bold"),
      legend.title = element_text(size = 9), 
      plot.subtitle = element_text(hjust = 1))
     
  
  ggsave(here("Output","TiltMeterData",WindrosePlotname))
  
 
   
  #TiltPhotoPlot <- as.ggplot(TiltPhoto)
  TiltPhotoPlot <- wrap_elements(
  panel = rasterGrob(as.raster(TiltPhoto),
                     width = unit(1, "npc"),
                     height = unit(0.5, "npc"))
)
  # TiltPhotoPlot <- wrap_elements(panel = rasterGrob(TiltPhoto, 
  #                                                   x = unit(1, "npc"), 
  #                                                   y = unit(1, "npc")))
  
  
 FeatherPlot2 <- FeatherPlot + 
   guides(shape = guide_legend(override.aes = list(size = 0.5)), 
          color = guide_legend(override.aes = list(size = 0.5))) +
   theme(legend.title = element_text(size = 4),
         legend.text = element_text(size = 4), 
         legend.key.size = unit(1, "mm"),       # size of boxes
    legend.key.height = unit(2, "mm"),     # height of each key row
    legend.key.width = unit(1.5, "mm"),      # width of each key column
    legend.spacing.y = unit(0.1, "mm"),    # vertical spacing between keys
    legend.spacing.x = unit(0.1, "mm"),    # horizontal spacing
          legend.box.background = element_rect(color = "black", size = 0.3), 
         axis.title.x =  element_blank(), 
         axis.text.x = element_blank(), 
         axis.ticks.x = element_blank())
         #legend.position = "none")
 WindrosePlot2 <- WindrosePlot + theme(plot.title = element_blank())

left_side  <- (FeatherPlot2 / TempPlot / TidePlot) + 
                plot_layout(axes = "collect")

right_side <- (WindrosePlot2 / plot_spacer()  / TiltPhotoPlot) 
  # plot_layout(heights = c(1,2), 
  #             widths = c(1,2))

combined <- (left_side | right_side) +
  plot_layout(widths = c(2,1)) +   # control relative widths
  plot_annotation(
    title = Sitename, 
    theme = theme(plot.title = element_text(hjust = 0.5, size = 16))
  ) &
  theme(text = element_text(face = "bold"))
  
  
  
  
  
 
# combined <- (FeatherPlot/TempPlot/TidePlot + plot_layout(width = 4, axes = "collect")) |
#   (WindrosePlot/TiltPhotoPlot +
# #combined <- (FeatherPlot + TempPlot + TidePlot + WindrosePlot + TiltPhotoPlot) +
#   plot_layout(width = 1)) +
#   plot_annotation(title = Sitename, 
#                   theme = theme(plot.title = element_text(size = 16))) &
#   
#   theme(text = element_text (face = "bold")) 
  

   ggsave(here::here("Output", "TiltMeterData", combinedplot))
 }



Velocity_TidePlot_fun(Tilt1DataHourly, FeatherPlotname = "Tilt1Velocity.png", Tilt1Temp, TempPlotname = "Tilt1Temp.png", Tilt1DataHourly, Sitename1 = "Site: Oki 2", AverageSpeed = "Average Speed: 1.74 cm/s", WindrosePlotname = "Windrose1.jpeg", Tilt1Photo, combinedplot = "Tilt1TideCombo.jpeg", Sitename = "Site: Oki 2")

 Velocity_TidePlot_fun(Tilt2DataHourly, FeatherPlotname = "Tilt2Velocity.png", Tilt2Temp, TempPlotname = "Tilt2Temp.png", Tilt2DataHourly, Sitename1 = "Site: Oki 3", AverageSpeed = "Average Speed: 0.80 cm/s", WindrosePlotname = "Windrose2.jpeg", Tilt2Photo, combinedplot = "Tilt2TideCombo.jpeg", Sitename = "Site: Oki 3")

 Velocity_TidePlot_fun(Tilt3DataHourly, FeatherPlotname = "Tilt3Velocity.png", Tilt3Temp, TempPlotname = "Tilt3Temp.png", Tilt3DataHourly, Sitename1 = "Site: Oki 4", AverageSpeed = "Average Speed: 3.53 cm/s", WindrosePlotname = "Windrose3.jpeg", Tilt3Photo, combinedplot = "Tilt3TideCombo.jpeg", Sitename = "Site: Oki 4")

 Velocity_TidePlot_fun(Tilt4DataHourly, FeatherPlotname = "Tilt4Velocity.png", Tilt4Temp, TempPlotname = "Tilt4Temp.png", Tilt4DataHourly, Sitename1 = "Site: Oki 7", AverageSpeed = "Average Speed: 6.27 cm/s", WindrosePlotname = "Windrose4.jpeg", Tilt4Photo, combinedplot = "Tilt4TideCombo.jpeg", Sitename = "Site: Oki 7")

 Velocity_TidePlot_fun(Tilt5DataHourly, FeatherPlotname = "Tilt5Velocity.png", Tilt5Temp, TempPlotname = "Tilt5Temp.png", Tilt5DataHourly, Sitename1 = "Site: Oki 9", AverageSpeed = "Average Speed: 2.28 cm/s", WindrosePlotname = "Windrose5.jpeg", Tilt5Photo, combinedplot = "Tilt5TideCombo.jpeg", Sitename = "Site: Oki 9")

 Velocity_TidePlot_fun(Tilt6DataHourly, FeatherPlotname = "Tilt6Velocity.png", Tilt6Temp, TempPlotname = "Tilt6Temp.png", Tilt6DataHourly, Sitename1 = "Site: Oki 11", AverageSpeed = "Average Speed: 0.80 cm/s", WindrosePlotname = "Windrose6.jpeg", Tilt6Photo, combinedplot = "Tilt6TideCombo.jpeg", Sitename = "Site: Oki 11")

 Velocity_TidePlot_fun(Tilt7DataHourly, FeatherPlotname = "Tilt7Velocity.png", Tilt7Temp, TempPlotname = "Tilt7Temp.png", Tilt7DataHourly, Sitename1 = "Site: Oki 15", AverageSpeed = "Average Speed: 1.45 cm/s", WindrosePlotname = "Windrose7.jpeg", Tilt7Photo, combinedplot = "Tilt7TideCombo.jpeg", Sitename = "Site: Oki 15")

 Velocity_TidePlot_fun(Tilt8DataHourly, FeatherPlotname = "Tilt8Velocity.png", Tilt8Temp, TempPlotname = "Tilt8Temp.png", Tilt8DataHourly, Sitename1 = "Site: Oki 14", AverageSpeed = "Average Speed: 5.43 cm/s", WindrosePlotname = "Windrose8.jpeg", Tilt8Photo, combinedplot = "Tilt8TideCombo.jpeg", Sitename = "Site: Oki 14")


```


# PlottingTidal Ellipses

### Prax Ellipse Function

This code is converted from Matlab Katie Smith Via Christina Comfort
First thing we have to do is make a function for the prax_ellipse

```{r}
prax_ellipse_fct <- function(u, v) {
  
  N <- length(u)
  
#Remove the mean over time, to leave jut the "pertubations"
  meanu <- mean(u)
  meanv <- mean(v)
  up <- u - meanu
  vp <- v - meanv
  
#This has to do with the mean variance (covariance matrix elements)
  # (1/N) * sum(vector * vector) is equivalent to N * var() but using N
  # Instead of N-1 fo rthe divisor in the variance/covariance calculation 
  # R's var() and cov() use N-1, so we calculate it manually as in Matlab. 
  
  c11 <- (1/N) * sum(up * up)
  c22 <- (1/N) * sum(vp *vp)
  c12 <- (1/N) * sum(up * vp)
  
  # This is the angle to rotate the vecotrs (in radians)
  # R's atan2 is the same as Matlab's
  
  phi <- 0.5 * atan2(2 *c12, c22 - c11) #+ pi/2
  phi_Deg <- (90 - phi * 180/pi) %% 360
  major_axis_angle <- phi_Deg
  
  #Eignevalues of a matrix of those values calculated above
  # R's eigen function returns a list with "values" (eigenvalues) and "vectors" (eigenvectors)
  
  cov_matrix <- matrix(c(c11, c12, c12, c22), nrow=2, byrow = TRUE)
  eig_result <- eigen(cov_matrix)

  #Extract eigenvalues
  # The **values** vector in the eig_result is already sorted in decreasing order by 
  lambda1 <- eig_result$values[1] #Major axis variance (largest eigenvalue)
  lambda2 <- eig_result$values[2] # Minor axis variance (smallest eigenvalue)
  
  #Here we turn the eigenvalues (which represetn the variance along the major and minor axes into varaiables which will make tidal ellipses)
  
  a <- sqrt(lambda1) #Major axis semi-length
  b <- sqrt(lambda2) # Minor axis semi-length
  
  #Create the parameter 't"
  
  t <- seq(0, 2 * pi, length.out = 100)
  
  # Calculate the ellipse coordinates (rotated and not centered/mean -subtracted yet)
  #Note: The orginal Matlab script has commented out adding 'meanu' to 'meanv', which would typically center teh ellipse on the mean velocity
  # Matlab's 'phi(1)' is used because 'phi' is an array of size4 1. In R, 'phi' is a scalar
  
  x_base <- a * cos(t)
  y_base <- b * sin(t)
  

  
 # x_rot <- a * cos(t) * cos(pi/2 - phi) - b * sin(t) * sin(pi/2 - phi)
 #y_rot <- a * cos(t) * sin(pi/2 - phi) + b * sin(t) * cos(pi/2 - phi)
  
  # Add the mean back to center the ellipse (if desired, based on the commented out Matlab line)
  # For consistency with the original script's output variable names, we'll use x_rot and y_rot
  # as 'x' and 'y', and include the mean adjustment as an option in the comments.
  # If you want the ellipse centered at the mean velocity:
  # x <- x_rot + meanu
  # y <- y_rot + meanv
   x_rot <- x_base * cos(phi) - y_base * sin(phi)
   y_rot <- x_base * sin(phi) + y_base * cos(phi)
  
  #The function will return the angle phi, and the coordinates x and y 
  return(list(phi = phi, x = x_rot, y = y_rot)) 
  
}

```

## Plotting Current Ellipses

We are again converting the matlab script and we are going to use the joined data to facet plot all the current ellipses together

Able to facet all the graphs into one, but now
```{r}
# Extract the velocity components

as.numeric(JoinedData[[8]]) #eastward
as.numeric(JoinedData[[7]]) # northward

#Run the Ellipse Funciton 


# JoinedData %>% mutate(results = prax_ellipse_fct(Velocity_N, Velocity_E))
# rot <- results$phi
# xx <- results$x
# yy <- results$y
# major_axis_angle <- (90 - rot * 180 / pi) %% 360

TiltEllipses <- JoinedData %>%
  group_by(TiltMeter) %>%
  summarise(
    ellipse = list({
      results <- prax_ellipse_fct(Velocity_N, Velocity_E) 
      phi_deg <- 90 - (results$phi * 180 / pi) 
  phi_deg <- ((phi_deg + 180) %% 360) -180
  
tibble(
  x = results$x, 
  y = results$y, 
phi_deg = phi_deg
)
}), 
.groups = "drop"
) %>%
  unnest(ellipse)

TiltEllipses <- TiltEllipses %>%
  group_by(TiltMeter) %>%
  mutate(angle_label = round(unique(phi_deg), 1))
view(TiltEllipses)

#Make a table so we can put all the tilt meter data together
# tibble(
#   Site = JoinedData$TiltMeter, 
#   x = results$x, 
#   y = results$y
# )

#Creating a dataframe to plot the ellipse coodinates 
# ellipse_data <- data.frame(x = xx, y = yy)

#Creating gg plot 

p <- ggplot(TiltEllipses, aes( x = x, y = y, color = TiltMeter)) +
  geom_path() +
  facet_wrap(~ paste0(TiltMeter, "\n", angle_label, "˚ cw from north"), scales = "fixed") +

  labs(

    x = 'u velocity (cm/s)',
    y = 'v velocity (cm/s)'
  ) +
# Enforce equal scale (axis equal equivalent)
  coord_fixed(ratio = 1) +
# Set the defined limits
  xlim(-7, 7) +
  ylim(-7, 7) +
  
  # Apply a clean theme
  theme_bw()

# Display the plot
print(p)

# p + 
#   geom_point(aes(x = mean(u), y = mean(v)), color = "red", size = 2) +
#   geom_segment(aes(x = 0, y = 0, xend = mean(u), yend = mean(v)), 
#                arrow = arrow(length = unit(0.2, "cm")), color = "red")
```

