---
title: "Tilt Meter Analysis"
author: "Ashley Carreiro"
format: html
editor: visual
---

# Load Libraries

```{r}
library(tidyverse)
library(here)
```

## Ignore: Creating New CSV Files

Trimming down the tilt meter data to one hour after deployment and one hour before retrieval of the tilt meters.

Making sure all data is captured at one minute intervals.

Note that all data has been adjusted for the Magnetic Declination of -5.85 West within the Domino app.

All data speed units are in cm/s

**This code does not need to be run again!**

```{r}
#| eval: false
#| echo: false
read_csv(here("Data/TiltMeterData/Tilt1_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% #Changing the headers int he dataframe
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% #Changing the date to character using lubridatde
  filter(between(DateTime, ymd_hms("2025-08-01 11:21:13"), ymd_hms("2025-08-14 08:45:13"))) %>% #Filtering between the dates that we want data from
  write_csv(file ="Data/TiltMeterData/Tilt1_Current.csv") #Creating new CSV File with filtered dates

read_csv(here("Data/TiltMeterData/Tillt2_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 13:00:42"), ymd_hms("2025-08-14 09:30:42"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt2_Current.csv")

read_csv(here("Data/TiltMeterData/Tilt3_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 14:45:50"), ymd_hms("2025-08-14T11:30:50"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt3_Current.csv")

#Tilt 4 was taking data every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt4_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 10:45:14"), ymd_hms("2025-08-14 12:45:04"))) %>% 
    group_by(DateTime = floor_date(DateTime, "min")) %>% #Averaging the data over 1 minute to match the other data
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(file ="Data/TiltMeterData/Tilt4_Current.csv")  

read_csv(here("Data/TiltMeterData/Tilt5_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 12:18:28"), ymd_hms("2025-08-15 12:30:28"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt5_Current.csv")

read_csv(here("Data/TiltMeterData/Tilt6_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 13:57:00"), ymd_hms("2025-08-15 11:30:00"))) %>% 
write_csv(file ="Data/TiltMeterData/Tilt6_Current.csv")

#Tilt 7 was taking readings every second
read_csv(here("Data/TiltMeterData/Tilt7_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 10:32:32"), ymd_hms("2025-08-15 08:46:00"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(file ="Data/TiltMeterData/Tilt7_Current.csv")

#Tilt 8 was taking readings every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt8_Current.csv")) %>%
 set_names(c("DateTime","Speed","Heading","Velocity_N","Velocity_E")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 12:06:06"), ymd_hms("2025-08-15 9:47:06"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(file ="Data/TiltMeterData/Tilt8_Current.csv")
```

Temperature Data

```{r}
#| eval: false
#| echo: false
 read_csv(here("Data/TiltMeterData/Tlt1_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% #Changing the headers int he dataframe
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% #Changing the date to character using lubridatde
  filter(between(DateTime, ymd_hms("2025-08-01 11:21:13"), ymd_hms("2025-08-14 08:45:13"))) %>%  #Filtering between the dates that we want data from
  write_csv(here("Data/TiltMeterData/Tilt1_Temperature.csv")) #Creating new CSV File with filtered dates

 read_csv(here("Data/TiltMeterData/Tilt2_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 13:00:42"), ymd_hms("2025-08-14 09:30:42"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt2_Temperature.csv"))
 
 read_csv(here("Data/TiltMeterData/Tilt3_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-01 14:45:50"), ymd_hms("2025-08-14T11:30:50"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt3_Temperature.csv"))
 
 #Tilt 4 was taking data every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt4_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 10:45:14"), ymd_hms("2025-08-14 12:45:04"))) %>% 
    group_by(DateTime = floor_date(DateTime, "min")) %>% #Averaging the data over 1 minute to match the other data
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(here("Data/TiltMeterData/Tilt4_Temperature.csv"))
 
read_csv(here("Data/TiltMeterData/Tilt5_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 12:18:28"), ymd_hms("2025-08-15 12:30:28"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt5_Temperature.csv"))

read_csv(here("Data/TiltMeterData/Tilt6_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-02 13:57:00"), ymd_hms("2025-08-15 11:30:00"))) %>% 
write_csv(here("Data/TiltMeterData/Tilt6_Temperature.csv"))

#Tilt 7 was taking readings every second
read_csv(here("Data/TiltMeterData/Tilt7_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 10:32:32"), ymd_hms("2025-08-15 08:46:00"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(here("Data/TiltMeterData/Tilt7_Temperature.csv"))

#Tilt 8 was taking readings every 20 seconds
read_csv(here("Data/TiltMeterData/Tilt8_Temperature.csv")) %>%
 set_names(c("DateTime","Temperature")) %>% 
  mutate(DateTime = lubridate::ymd_hms(DateTime)) %>% 
  filter(between(DateTime, ymd_hms("2025-08-03 12:06:06"), ymd_hms("2025-08-15 9:47:06"))) %>% 
      group_by(DateTime = floor_date(DateTime, "min")) %>% 
   summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
write_csv(here("Data/TiltMeterData/Tilt8_Temperature.csv"))

```

Doing the same for the tide data

Tide data was downloaded from: <https://www.gsi.go.jp/kanshi/tide_data_21_e.html>

```{r}
#| eval: false
#| echo: false
read_table(here("Data/TiltMeterData/21_2021-2025_hour.txt"), col_names = FALSE) %>%
select(-1) %>% #Drops the first column
set_names(c("Date", "Hour", "Sealevel_mm", "DatumConstant_mm", "FixedPointHeight_mm")) %>%
 mutate(Date = lubridate::ymd(Date)) %>% 
  filter(between(Date, ymd("2025-08-01"), ymd("2025-08-15"))) %>%
  
write_csv(here("Data/TiltMeterData/TideData_2025.08.csv"))

```

## Load Data

Load current, temperature, and tide data from new csv files created above.

```{r}
#Load Current Data
Tilt1Data <- read_csv(here("Data", "TiltMeterData", "Tilt1_Current.csv"))
Tilt2Data <- read_csv(here("Data", "TiltMeterData", "Tilt2_Current.csv"))
Tilt3Data <- read_csv(here("Data","TiltMeterData", "Tilt3_Current.csv"))
Tilt4Data <- read_csv(here("Data","TiltMeterData","Tilt4_Current.csv"))
Tilt5Data <- read_csv(here("Data","TiltMeterData", "Tilt5_Current.csv"))
Tilt6Data <- read_csv(here("Data","TiltMeterData", "Tilt6_Current.csv"))
Tilt7Data <- read_csv(here("Data","TiltMeterData", "Tilt7_Current.csv"))
Tilt8Data <- read_csv(here("Data","TiltMeterData", "Tilt8_Current.csv"))

#Load Temp Data

Tilt1Temp <- read_csv(here("Data/TiltMeterData/Tilt1_Temperature.csv"))
Tilt2Temp <- read_csv(here("Data/TiltMeterData/Tilt2_Temperature.csv"))
Tilt3Temp <- read_csv(here("Data/TiltMeterData/Tilt3_Temperature.csv"))
Tilt4Temp <- read_csv(here("Data/TiltMeterData/Tilt4_Temperature.csv"))
Tilt5Temp <- read_csv(here("Data/TiltMeterData/Tilt5_Temperature.csv"))
Tilt6Temp <- read_csv(here("Data/TiltMeterData/Tilt6_Temperature.csv"))
Tilt7Temp <- read_csv(here("Data/TiltMeterData/Tilt7_Temperature.csv"))
Tilt8Temp <- read_csv(here("Data/TiltMeterData/Tilt8_Temperature.csv"))

#Load Tide Data

TideData <- read_csv(here("Data/TiltMeterData/TideData_2025.08.csv"))


```

# Summary Stats 

First, join all the Current Data

```{r}
JoinedData <- rbind(Tilt1DataClean, Tilt2DataClean, Tilt3DataClean, Tilt4DataClean, Tilt5DataClean,  Tilt6DataClean,  Tilt7DataClean, Tilt8DataClean)

```

Visualize Summary stats to see where to filter data from **Don't Need to Run Everytime**

```{r}
summary(JoinedData$Speed)
lower_bound <- quantile(JoinedData$Speed, 0.001)
Upper_bound <- quantile(JoinedData$Speed, 0.999)
Upper_bound#Upper bound is 25.01416 we will filter data from this number 

boxplot.stats(JoinedData$Speed_m)

sum(JoinedData$Speed >= .5) #Counting the number of points above .5
# = 5

sum(JoinedData$Speed> 25.01416) #Number of points above the upper bound, 0.25, 145 points

table(cut(JoinedData$Speed_m , seq(0, .25,.05), include.lowest = TRUE)) ##Showing me the different bins and how many data points I have in each one

Means <- JoinedData %>%
  group_by(TiltMeter) %>%
  summarise(mean(Speed_m))
Means #Giving the average speed m/s of each tilt meter

#Visualize the data 

ggplot(JoinedData, aes(x = DateTime, y = Speed_m)) +
  geom_point() +
   facet_wrap(~TiltMeter)
```

# Organizing the data

### Tilt Meter Data

Creating a function to organize all the current data

-   Separating Data

-   

```{r}
Tilt_Data_fun <- function(df, tilt_name){  #Function to add a tilt name column and separate data and time column 
df %>%
    
separate(col = DateTime, #Choose the column that you want separated
             into = c("Date", "Time"), # Separate it into two columns "Data" and "Time"
             sep = " ", #What are the two separated by 
             remove = FALSE) %>% #Keep the orginial DateTime Column    
    
mutate( #Set datetime,a dding in the UTC timezone so wer are removing taht
       Date = lubridate::ymd(Date), #Set Date
       Time = lubridate::hms(Time), #Set Time
       Speed_m = Speed * 0.01, #Change cm/s -> m/s
       VelocityN_m = Velocity_N * 0.01, #Change cm/s -> m/s
       VelocityE_m = Velocity_E* 0.01, #Change cm/s -> m/s
   
       #Double check that there are no negative headings, can do 360 + negative
       
       TiltMeter = tilt_name) %>%
    
    select(TiltMeter, everything()) #Moving TiltMeter# Column to be the first column
    
}

Tilt1DataClean <- Tilt_Data_fun(Tilt1Data, "Tilt1")
Tilt2DataClean <- Tilt_Data_fun(Tilt2Data, "Tilt2")
Tilt3DataClean <- Tilt_Data_fun(Tilt3Data, "Tilt3")
Tilt4DataClean <- Tilt_Data_fun(Tilt4Data, "Tilt4")
Tilt5DataClean <- Tilt_Data_fun(Tilt5Data, "Tilt5")
Tilt6DataClean <- Tilt_Data_fun(Tilt6Data, "Tilt6")
Tilt7DataClean <- Tilt_Data_fun(Tilt7Data, "Tilt7")
Tilt8DataClean <- Tilt_Data_fun(Tilt8Data, "Tilt8")
```

Now we will average everything by hour and create a new data frame **hourly**

This dataframe is in cm/s

Filtering out the 99.9% upper bound which is anything above 25.01416 cm/s

```{r}
HourlyTilt_Data_fun <- function(df){  #Function to add a tilt name column and separate data and time column 
 df %>% 
    
    group_by(DateTime = floor_date(DateTime, "hour")) %>% #New datetime that is rounded to the start of each hour
  summarise(across(where(is.numeric), mean, na.rm = TRUE),
    .groups = "drop") %>%
    filter(Speed <=  25.01416) 
}

Tilt1DataHourly <- HourlyTilt_Data_fun(Tilt1Data)
Tilt2DataHourly <- HourlyTilt_Data_fun(Tilt2Data)
Tilt3DataHourly <- HourlyTilt_Data_fun(Tilt3Data)
Tilt4DataHourly <- HourlyTilt_Data_fun(Tilt4Data)
Tilt5DataHourly <- HourlyTilt_Data_fun(Tilt5Data)
Tilt6DataHourly <- HourlyTilt_Data_fun(Tilt6Data)
Tilt7DataHourly <- HourlyTilt_Data_fun(Tilt7Data)
Tilt8DataHourly <- HourlyTilt_Data_fun(Tilt8Data)

```

### Temperature Data

```{r}
Tilt_Temp_fun <- function(df){ 
df %>%
    
separate(col = DateTime, #Choose the column that you want separated
         into = c("Date", "Time"), # Separate it into two columns "Data" and "Time"
         sep = " ", #What are the two separated by 
         remove = FALSE) %>% #Keep the orginial DateTime Column    
  mutate(DateTime = lubridate::ymd_hms(DateTime), #Set datetime
         Date = lubridate::ymd(Date), #Set Date
         Time = lubridate::hms(Time))
}

Tilt1TempData <- Tilt_Temp_fun(Tilt1Temp)
Tilt2TempData <- Tilt_Temp_fun(Tilt2Temp)
Tilt3TempData <- Tilt_Temp_fun(Tilt3Temp)
Tilt4TempData <- Tilt_Temp_fun(Tilt4Temp)
Tilt5TempData <- Tilt_Temp_fun(Tilt5Temp)
Tilt6TempData <- Tilt_Temp_fun(Tilt6Temp)
Tilt7TempData <- Tilt_Temp_fun(Tilt7Temp)
Tilt8TempData <- Tilt_Temp_fun(Tilt8Temp)
```

### Tide Data

```{r}
 TideDataHourly <- TideData %>%
   mutate(DateTime = ymd_hms(paste(Date, Hour, "00:00")), #Combining Data and time into one column 
          Sealevel_m = Sealevel_mm / 1000) %>%
    select(DateTime, everything())

```

# Plotting the Data

We are going to make a big function to make a plot of the following

-   Feather Plot. \| Wind Rose Plot

-   Temperature Plot \| Elipses Plot?

-   Tide Plot \| Photo of Tilt Meter?

```{r}
Velocity_TidePlot_fun <- function(Tiltdf, FeatherPlotname, Tempdf, TempPlotname, Tiltdf2, WindrosePlotname, combinedplot) {
  
time_limits <- range(Tiltdf$DateTime, na.rm = TRUE)

FeatherPlot <- ggplot(Tiltdf) +
 
  geom_segment(aes( x = DateTime, y = 0, 
                    xend = DateTime, 
                    yend = Velocity_N #Makes the arrows point up/down
                    ), 
               color = "blue",
               arrow = arrow(length = unit(0.2, "cm")),
               linewidth = 0.5,
               alpha = 0.5) +
  
  geom_segment(aes( x = DateTime, y = 0, 
                    xend = DateTime  , #Makes the arrows lean left/right
                    yend = Velocity_E #Makes the arrows point up/down
                    ), 
               color = "red",
               arrow = arrow(length = unit(0.2, "cm")),
               linewidth = 0.5, 
               alpha = 0.5)+
  
  scale_x_datetime(limits =  time_limits) +
  labs(
     
    x = "Time", 
    y = "Velocity (cm/s)") +
  theme_minimal()  
 
 ggsave(here("Output","TiltMeterData",FeatherPlotname))
 
 
 TempPlot <- ggplot(Tempdf, aes(DateTime, Temperature)) +
  geom_point()
  
ggsave(here("Output", "TiltMeterData", TempPlotname))
 


TidePlot <- ggplot(TideDataHourly, aes(DateTime, Sealevel_m)) +
   geom_line(color = "steelblue") +
   scale_x_datetime(limits = time_limits) 
   
  
 
 WindrosePlot <- ggwindrose(
    speed = Tiltdf2$Speed, 
    direction = Tiltdf2$Heading,
    n_directions = 8,
    speed_cuts = seq(0, 25, 5), Inf) +
    #stack_revers = TRUE) + #If we want the higher speeds on the outside of the plot
  

    scale_fill_manual(
      values = c( "skyblue","orange" ,"green2", "blue","red", "darkred"),
      drop = FALSE
    ) +
    
    labs(fill = "Current Speed (cm/s)") +
      theme_minimal() +
    theme(axis.title = element_blank(), 
          text = element_text(face = "bold"))
     
  
  ggsave(here("Output","TiltMeterData",WindrosePlotname))
 
combined <- plot_grid(FeatherPlot, TempPlot, TidePlot, WindrosePlot, ncol = 1, align = "v", rel_heights =  c(2,1))

   ggsave(here::here("Output", "TiltMeterData", combinedplot))
 }



Velocity_TidePlot_fun(Tilt1DataHourly, FeatherPlotname = "Tilt1Velocity.png", Tilt1Temp, TempPlotname = "Tilt1Temp.png", Tilt1DataHourly, WindrosePlotname = "Windrose1.jpeg", combinedplot = "Tilt1TideCombo.png")

Velocity_TidePlot_fun(Tilt2DataHourly, FeatherPlotname = "Tilt2Velocity.png", Tilt2Temp, TempPlotname = "Tilt2Temp.png", Tilt2DataHourly, WindrosePlotname = "Windrose2.jpeg", combinedplot = "Tilt2TideCombo.png")

Velocity_TidePlot_fun(Tilt3DataHourly, FeatherPlotname = "Tilt3Velocity.png", Tilt3Temp, TempPlotname = "Tilt3Temp.png", Tilt3DataHourly, WindrosePlotname = "Windrose3.jpeg", combinedplot = "Tilt3TideCombo.png")

Velocity_TidePlot_fun(Tilt4DataHourly, FeatherPlotname = "Tilt4Velocity.png", Tilt4Temp, TempPlotname = "Tilt4Temp.png", Tilt4DataHourly, WindrosePlotname = "Windrose4.jpeg", combinedplot = "Tilt4TideCombo.png")

Velocity_TidePlot_fun(Tilt5DataHourly, FeatherPlotname = "Tilt5Velocity.png", Tilt5Temp, TempPlotname = "Tilt5Temp.png", Tilt5DataHourly, WindrosePlotname = "Windrose5.jpeg", combinedplot = "Tilt5TideCombo.png")
```

